---
title: "Multiple Linear Regression"
subtitle: "MATH 325 - Brother Saunders"
author: "Andrew Seaman"
date: "02-17-2026"
output: 
  
  html_document:
    theme: space
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 6
editor_options: 
  chunk_output_type: console
---

<!-- ————————————————————————————————————————————————————————————
INSTRUCTIONS
—————————————————————————————————————————————————————————————— -->
```{r directions, echo=FALSE}
# Come up with a simple linear regression of your own creation. 

# * Write the Background and clearly state your question. Then collect data and perform the linear regression analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.
 
# * You should collect both the quantitative response variable and the quantitative explanatory variable yourself. This can be data you collect from 
#     - the internet (like weather data, Facebook data, or sports data) 
#     - your own experiment 
#     - your own observational study
# * However,     
#     - The data should not be a single pre-existing dataset. It must be something that you piece together yourself.
 
# * You need at least $n=3$ $(X,Y)$ pairs to perform the regression. However, your regression will only be useful if you can get around $n=10$ or more observations.

# You can use either:

#   someName <- read.csv("../Data/YourDataFileName.csv", header=TRUE)

# or:
 
#   library(readr)
#   someName <- read_csv("../Data/YourDataFileName.csv")
 
 
# Do not forget to run: "Session -> Set Working Directory -> To Source file location"
```


## Introduction {.tabset .tabset-pills .tabset-fade}

<!--———————————————————————————————————————————————————————————
BACKGROUND
————————————————————————————————————————————————————————————-->
### Background

**Data Collection with Apple Devices**

The iPhone [^1] (*June 29, 2007*) in tandem with the Apple Watch [^2](*April 24,2015*) record numerous health metrics for the user/wearer made accessible in the IOS Health App [^3] (*September 17, 2014*). Per Apple's promotions, users are intended to use these metrics to set goals and train more effectively. Two intriguing metric to avid and casual health conscious users alike is daily total step count and average heart rate. Measurements are more accurate for individuals who use an Apple Watch because wearable devices tend to be on the user's person more consistently than an iPhone especially during exercise when a lot of valuable data is collected. In particular, Heart Rate can't be recorded with an iPhone alone.

**Subject's Device Use History**

The subject of this study, who is also the author, first began collecting health metrics in the late 2010s. Below is a timeline of device acquisition an subsequent use:

1. Sep 2017 - iPhone SE [^4]
2. Jun 2019 - iPhone 7 [^5]
3. Oct 2022 - iPhone 12 Pro [^6]
4. Jul 2024 - Apple Watch 1 [^7]
5. Jan 2025 - iPhone 16 Pro Max [^8]
6. Feb 2025 - Apple Watch 10 [^9]


<!--———————————————————————————————————————————————————————————
SETUP
————————————————————————————————————————————————————————————-->

### Setup

First, several code libraries were used to perform the statistical tests.

<!-- LIBRARIES -->
<details>
<summary><strong>Libraries</strong></summary>
These **libraries** are used to reproduce this analysis.
```{r libraries, class.source = 'fold-show', message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(mosaic)
library(car)
library(pander)
library(readr) # Used to import the data
library(scales) # Needed? ChatGPT gave it to me when helping me write the big plotting function (jitter-box-line-plot 1 and 2)
library(directlabels)
#library(lubridate) # For date extraction in the 2nd plot of the Graphics section
```
</details>

</BR>

<!-- DATA -->
Next, the author's health data was imported and cleaned to improve readability.

<details>
<summary><strong>Data</strong></summary>
This **data** is imported locally and the date field data type is set to a date type.
```{r dataset, class.source = 'fold-show'}
# Data import
# file_path <- "../../Data/iosHealth/health_export_2026-01-12.csv"
# health <- read_csv(file_path, col_types = cols(Date = col_date(format = "%m/%d/%y")))

health <- read_csv("../../Data/iosHealth/IOS_health_2017-2026.csv",
    col_types = cols(Date = col_date(format = "%m/%d/%y")))

#  [1] "Date"                                        
#  [2] "Cycling Distance (sum)"                      
#  [3] "Exercise Minutes (sum)"                      
#  [4] "Lean Body Mass (most recent)"                
#  [5] "Walking/Running Distance (sum)"              
#  [6] "Walking HR Avg (average)"                    
#  [7] "Mindful Minutes (sum)"                       
#  [8] "Body Temperature (Average) (average)"        
#  [9] "Basal Energy (sum)"                          
# [10] "Blood Glucose (Most Recent) (most recent)"   
# [11] "Heart Rate (average)"                        
# [12] "Flights Climbed (sum)"                       
# [13] "Headphone Audio (average)"                   
# [14] "Respiratory Rate (average)"                  
# [15] "Body Fat % (most recent)"                    
# [16] "Environmental Audio (average)"               
# [17] "Swimming Distance (sum)"                     
# [18] "Blood Oxygen (average)"                      
# [19] "Body Temperature (Most Recent) (most recent)"
# [20] "Sleep (sum)"                                 
# [21] "BP Diastolic (Average) (average)"            
# [22] "VO2 Max (most recent)"                       
# [23] "BP Systolic (Most Recent) (most recent)"     
# [24] "BP Systolic (Average) (average)"             
# [25] "Blood Glucose (Average) (average)"           
# [26] "Sleep (Core) (sum)"                          
# [27] "Total Energy (sum)"                          
# [28] "HRV (SDNN) (average)"                        
# [29] "BP Diastolic (Most Recent) (most recent)"    
# [30] "Steps (sum)"                                 
# [31] "Weight (Most Recent) (most recent)"          
# [32] "Resting Heart Rate (average)"                
# [33] "Sleep (REM) (sum)"                           
# [34] "Sleep (Asleep) (sum)"                        
# [35] "Sleep (Deep) (sum)"  

```

<!-- DATA CLEANING -->

Lastly, the original **column names** were cleaned up.

<!-- Clean Data -->
```{r rename-fields, class.source = 'fold-show'}
# > Select columns, filter by year, and rename fields
health <- health %>% 
  rename(
    HeartRate = `Heart Rate (average)`,
    Steps = `Steps (sum)`
  ) %>% 
  select(Date, HeartRate, Steps) %>% 
# > Only use data from 2023-2026
  filter(year(Date) >= 2023)
```

<!-- Add `Year` Column -->
  <!-- Extract year value from `Date` column. -->
```{r add-date-col}
health <- mutate(health, Year = year(Date))
```

<!-- Add Dummy Column: "Is 2023?" -->
```{r add_dummy_col_2023}
health <- health %>% 
  mutate(
    is_2023 = ifelse(Year == 2023, 1, 0))
```
</details>

### Hide

</BR>


<!--———————————————————————————————————————————————————————————
ANALYSIS
————————————————————————————————————————————————————————————-->

## Analysis

<!-- TEMPORARY HEADER. Used to collapse hypotheses section. REMOVE UPON PUBLISH. -->
<!-- ### *Hypothesis (temp)* -->

> #### Hypothesis

This analysis aims to model the actual daily **average heart rate** (*Avg Heart Rate*) of an individual according to their recorded daily **total steps** (*Total Steps*). Mathematically,

<!-- TeX commands taken from:
https://byuistats.github.io/Statistics-Notebook/Analyses/Linear%20Regression/Examples/BodyWeightSLR.html# -->

$$
\underbrace{Y_i}_\text{Avg Heart Rate} = \beta_0 + \beta_1 \underbrace{X_i}_\text{Total Steps} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2)
$$

The hypotheses below closely observe the **true slope** of the regression model, $\beta_1$. If $\beta_1$ is 0, then we're confident that between *Avg Heart Rate* and reported *Total Steps* there is not a meaningful relationship. The linear regression model would be considered neither positive nor negative but *dead* where the one variables does not impact the other.

$$
H_0: \beta_1 = 0 \\  
H_a: \beta_1 \neq 0 \\
\alpha = 0.1
$$


<!-- TEMPORARY HEADER. Used to collapse hypotheses section. REMOVE UPON PUBLISH. -->
<!-- ### *Graphics (temp)* -->

> #### Graphics

First, we'll plot each variable independently and look at the variance year-over-year (YoY) before comparing the variables against each other looking for covariance where one variable affects the other.

---

<!-- Plot Setup: Base plot Variable: Daily Total {var} Over Time (5 yrs) -->
```{r fig1-help}
# 1) Function: df of means by Year
df_means <- function(df, var) {
  df %>%
    filter(!is.na({{ var }}), !is.na(Year)) %>%
    group_by(Year) %>%
    summarize(mean = mean({{ var }}, na.rm = TRUE), .groups = "drop")
}



# 2) Function: counts of non-missing points by Year
df_counts <- function(df, var) {
  df %>%
    filter(!is.na({{ var }}), !is.na(Year)) %>%
    count(Year, name = "Points")
}



# 3) Function: Jitter Box Line Plot
# > Function start and parameters       ylimits: optional parameter
jitter_box_line_plot <- function(df, var, title, subtitle, ylab, xlab) {
# > Clean dataframe
  df_clean <- df %>%
    filter(!is.na({{ var }}), !is.na(Year))
# > Order years
  year_levels <- sort(unique(df_clean$Year))
# > Base plot
  ggplot(df_clean, aes(x = factor(Year), y = {{ var }}, color = factor(Year))) +
#   > Boxplot
    geom_boxplot(outlier.alpha = 0) +
#   > Jitter plot
    geom_jitter(alpha = 0.2, width = 0.2) +
#   > Line of means
    geom_line(
      data = df_means(df, {{ var }}),
      aes(x = factor(Year), y = mean, group = 1),
      inherit.aes = FALSE
    ) +
#   > Specify upper y-axis max limit    
    # scale_y_continuous(limits = {{ylimits}})+
#   > Counts
    geom_label(
      data = df_counts(df, {{ var }}),
      aes(x = factor(Year), y = Inf, label = Points),
      vjust = 1.5,
      show.legend = FALSE,
      inherit.aes = FALSE
    ) +
#   > Label for counts
    annotate(
      "text",
      x = as.character(min(year_levels)),  # leftmost year (factor-safe)
      y = Inf,
      label = "Points:",
      vjust = 2.2,
      hjust = 1.5
    ) +
#   > Labels
    labs(title = title, subtitle = subtitle, y = ylab, x = xlab) +
#   > Color scale
    scale_color_brewer(palette = "Set1", name = "Year") +
#   > Base theme
    theme_bw() +
#   > Custom theme
    theme(
      legend.position = "none",
      plot.margin = margin(10, 30, 10, 10)
    ) +
#   > ??
    coord_cartesian(clip = "off")
}
```

Variable: **`Heart Rate`**

In the plot below, 2026 shows three things of note:

1. High right skew or greater spread in the upper 50% of the data about twice that of other years.
2. The median is higher than for others years.
3. The mean is much further away (above) from the median than for other years.

Additionally, because the subject had sporadic access to heart rate measuring devices until mid-2024 data collection is scarce in 2023.

<!-- Plot: Boxplot + Scatterplot: Daily {Avg Heart rate} Over Time (5 yrs) -->
```{r fig2-avg-heart-rate, warning=FALSE, message=FALSE}
jitter_box_line_plot(
  health,
  HeartRate,
  # c(50,175),
  "Daily Average Heart Rate YoY (4 yrs)",
  "Black line marks the means; Densest: 2024-2025",
  "Heart Rate",
  "Year"
)
```

Variable: **`Steps`**

In the the plot below, there is a noticeable spike in the middle of 2024. The individual was employed to do outdoor work and averaged **≥ 20,000** steps almost daily for a couple of months from July-August.

<!-- Plot: Boxplot + Scatterplot: Daily {Total Steps} Year-Over-Year (5 yrs) -->
```{r fig2-total-steps, warning=FALSE, message=FALSE}
jitter_box_line_plot(
  health,
  Steps,
  # c(0,30000),
  "Daily Total Steps YoY (4 yrs)",
  "Black line marks the means; Densest: 2023-2025; Spike: mid-to-late 2024",
  "Steps",
  "Year"
)
```

Although the plots above hint at a time trend in the data, returning to the primary focus of this analysis the plot below models the linear relationship between heart rate and steps. We see a positive correlation (blue line) that passes through the intersection of $\bar{x}$ and $\bar{y}$ (red lines) as expected.

<!-- Plot: Comparing both variables -->
```{r plot_covariance, warning=FALSE, message=FALSE}
# ATTEMPT 1
# ——————————————————————————————————————————————


# # NOTE FOR BROTHER SAUNDERS (2/17/26):
# #   - I figured out why the lm (blue line) didn't pass through xbar and y bar (red lines intersection)
# #   - The means, medians, and lm were all calculated using different subsets.
# #   - If a row had Heart Rate but not steps the mean and median line would include that row.
# #   - However, lm requires both Heart Rate and Steps to have a value (or be finte) for plotting
# 
# 
# # Clean df where both variables have value in the same row
# df_finite <- health %>% 
#   filter(is.finite(HeartRate), is.finite(Steps))
# 
# df_2023 <- df_finite %>% 
#   filter(is_2023 == 1)
# 
# # Line colors
# meanR_medianG <- c("red","green")
# 
# 
# 
# # > Plot Start
# ggplot(df_finite, aes(x = HeartRate, y = Steps, color=as.factor(Year))) +
# # > Scatterplot
#   geom_point(pch = 19, size = 0.5) +
# # > Base theme
#   theme_bw() +
# # > Custom theme
#   theme(
# #   > Black background and white border
#     panel.background = element_rect(fill = "black"),
#     plot.background  = element_rect(fill = "black"),
#     panel.border     = element_rect(color = "white"),
# #   > White text
#     text = element_text(color = "white"),
#     axis.text = element_text(color = "white"),
#     axis.title = element_text(color = "white"),
# #   > Grey grid lines
#     panel.grid.major = element_line(color = "grey30"),
#     panel.grid.minor = element_line(color = "grey20"),
#     legend.position = "none"
#   )+
# # > Custom titles/labels
#   labs(
#     title = "More daily steps predict higher daily average heart rate",
#     subtitle = "Green: median; Red: mean; Blue: Linear Model",
#     x = "Daily Avg Heart Rate",
#     y = "Daily Total Steps")+
# # > Vertical Lines
#   geom_vline(
#     xintercept = c(mean(df_finite$HeartRate), median(df_finite$HeartRate)),
#     color = meanR_medianG,
#     linetype = "dashed")+
# # > Horizontal Lines
#   geom_hline(
#     yintercept = c(mean(df_finite$Steps), median(df_finite$Steps)),
#     color = meanR_medianG,
#     linetype = "dashed")+
# # > Linear Model Line: ALL
#   geom_smooth(
#     method = "lm",
#     color = "grey50",
#     fill = "grey50",
#     se = FALSE)+
# # > Linear Model Line: 2023
#   geom_smooth(
#     aes(data = df_2023, x = HeartRate, y = Steps),
#     method = "lm",
#     se = TRUE
#   )
```


```{r scatterplot_bivarite_by_year, message=FALSE, warning=FALSE}
# ATTEMPT 2 (works!)
# ——————————————————————————————————————————————
# Clean df where both variables have value in the same row
df_finite <- health %>% 
  filter(is.finite(HeartRate), is.finite(Steps), !is.na(Year))

df_2023 <- df_finite %>% 
  filter(Year == 2023)   # (or is_2023 == 1 if that column exists)

df_not_2023 <- df_finite %>%
  filter(Year != 2023)

meanR_medianG <- c("red", "green")



ggplot(df_finite, aes(x = HeartRate, y = Steps, color = as.factor(Year))) +
  geom_point(data = df_2023, color = "#F8766D", pch = 1, size = 2, stroke = 1) +
  geom_point(data = df_2023, color = "#F8766D", pch = 19, size = 0.5) +
  geom_point(color = "grey60", pch = 19, size = 0.5, alpha = .5) +
  theme_bw() +
  theme(
    panel.background = element_rect(fill = "black"),
    plot.background  = element_rect(fill = "black"),
    panel.border     = element_rect(color = "white"),
    text = element_text(color = "white"),
    axis.text = element_text(color = "white"),
    axis.title = element_text(color = "white"),
    panel.grid.major = element_line(color = "grey30"),
    panel.grid.minor = element_line(color = "grey20"),
    legend.position = "none"
  ) +
  labs(
    title = "More daily steps predict higher daily average heart rate",
    subtitle = "Green: median; Red: mean; Blue: 2023 Linear Model (others grey, no CI)",
    x = "Daily Avg Heart Rate",
    y = "Daily Total Steps"
  ) +
  geom_vline(
    xintercept = c(mean(df_finite$HeartRate), median(df_finite$HeartRate)),
    color = meanR_medianG,
    linetype = "dashed"
  ) +
  geom_hline(
    yintercept = c(mean(df_finite$Steps), median(df_finite$Steps)),
    color = meanR_medianG,
    linetype = "dashed"
  ) +

  # ---- Smooth lines for 2024–2026: gray, NO cloud, one per year ----
  geom_smooth(
    data = df_not_2023,
    aes(x = HeartRate, y = Steps, group = Year),
    method = "lm",
    se = FALSE,
    inherit.aes = FALSE,
    color = "grey60",
    linetype = "dashed"
  ) +
  
  # geom_smooth(method = "lm", se = FALSE, linetype = "dashed")+

  # ---- Smooth line for 2023: colored + cloud ----
  geom_smooth(
    data = df_2023,
    aes(x = HeartRate, y = Steps),
    method = "lm",
    se = TRUE,
    inherit.aes = FALSE,
    color = "#F8766D",
    fill = "#F8766D",
    alpha = 0.15
  )+
  geom_dl(label = df_finite$Year,
          method = list(
            "first.points",
            hjust = c(1.1,0,.8,1.3),
            vjust = c(3.5,-11,-3.5,-9.5)
          )
  )

  # scale_color_manual(color = c("deepskyblue", "grey60", "grey60", "grey60"))
```



```{r}
# 
```







## Test

> Simple Linear Regression Model Summary

Below we read from the summary of the linear regression model and see that the intercept (75.83) is not very interesting because it does not make much sense. It essentially says, "On the day you die you will walk 75.83 steps and your heart rate will be zero." However, the p-value is highly significant being much smaller than our hypothesized $\alpha$ of .05.

```{r}
mylm <- lm(HeartRate ~ Steps, data = health)
summary(mylm) %>% pander()
```

> Diagnostic Plots

**Residuals vs Fitted**

On the left we see that when Heart Rate is <90 BPM the variance is constant but >90 BPM and the data trends downward indicating that higher heart rate may not correlate with as much steps as a lower heart rate. Perhaps other activities such as running which required fewer steps gets the heart rate up and creates the trend we see.

**Q-Q Residuals**

At first glance one might think the data is mostly normal until the upper tail deviates from the theoretical line however this is false. With normality it is an all or nothing determination. If you plotted the data as a distribution curve it would have a tail like we see here. The only way to "make it normal" would be to cut off it's right tail but you can't do that because that would misrepresent the real trends observed in the data. 

Therefore, the data does <strong style="color: red;">not appear to be normal</strong> and fails the test.

**Plot 3**

There appears to be a trend in the data with tails in the lower left and in upper right as well as a de-clustering in the far right showing a possible downward drift in the data. This shows these variables are not independent.

```{r}
par(mfrow=c(1,3))
plot(mylm,which=1:2)
plot(mylm$residuals)
```



<!--———————————————————————————————————————————————————————————
IDEAS
————————————————————————————————————————————————————————————-->

<!-- TEMPORARY HEADER. Used to collapse hypotheses section. REMOVE UPON PUBLISH. -->
<!-- # *Ideas (temp)* -->

<!-- 1. Build facets of `heartRate` and `steps` independently for each year to best flesh out the height and range of each variable by year. -->




<!--———————————————————————————————————————————————————————————
EXPLORATION 
————————————————————————————————————————————————————————————-->


<!--———————————————————————————————————————————————————————————
FOOTNOTES
————————————————————————————————————————————————————————————-->

> Footnotes

[^1]: [iPhone](https://www.apple.com/iphone/)
[^2]: [Apple Watch](https://www.apple.com/watch/)
[^3]: [iOS Health App](https://www.apple.com/health/)
[^4]: [iPhone SE](https://support.apple.com/en-us/112005)
[^5]: [iPhone](https://support.apple.com/en-us/111943)
[^6]: [iPhone 12 Pro](https://support.apple.com/en-us/111875)
[^7]: [Apple Watch 1](https://support.apple.com/en-us/111985)
[^8]: [iPhone 16 Pro Max](https://support.apple.com/en-us/121032)
[^9]: [Apple Watch Series 10](https://support.apple.com/en-us/121202)

<!-- ************ -->
<!-- FILE END -->
<!-- ************ -->