---
title: "Simple Linear Regression"
output: 
  html_document:
    theme: space
    code_folding: hide
    toc: true
    toc_float: true
    toc_depth: 3
editor_options: 
  chunk_output_type: console
---

<!-- ————————————————————————————————————————————————————————————
DIRECTIONS
—————————————————————————————————————————————————————————————— -->
```{r directions, echo=FALSE}
# DIRECTIONS:
# 
# Come up with a simple linear regression of your own creation. 
# 
# * Write the Background and clearly state your question. Then collect data and perform the linear regression analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.
# 
# * You should collect both the quantitative response variable and the quantitative explanatory variable yourself. This can be data you collect from 
#     - the internet (like weather data, Facebook data, or sports data) 
#     - your own experiment 
#     - your own observational study
# * However,     
#     - The data should not be a single pre-existing dataset. It must be something that you piece together yourself.
# 
# * You need at least $n=3$ $(X,Y)$ pairs to perform the regression. However, your regression will only be useful if you can get around $n=10$ or more observations.
# 
# You can use either:
# 
#   someName <- read.csv("../Data/YourDataFileName.csv", header=TRUE)
# 
# or:
# 
#   library(readr)
#   someName <- read_csv("../Data/YourDataFileName.csv")
# 
# 
# Do not forget to run: "Session -> Set Working Directory -> To Source file location"
```

<!--———————————————————————————————————————————————————————————
SETUP
————————————————————————————————————————————————————————————-->

## Setup {.tabset .tabset-pills .tabset-fade}

### Show

For this analysis several code libraries were relied on to perform the statistical tests. Next, the author's health data was imported and cleaned up to upgrade the readability of the data.

<!-- LIBRARIES -->
<details>
<summary><strong>Libraries</strong></summary>

These **libraries** are used to reproduce this analysis.

```{r library-import, class.source = 'fold-show', message=FALSE, warning=FALSE}
# Libraries
library(tidyverse)
library(mosaic)
library(car)
library(pander)
library(readr) # Used to import the data
#library(lubridate) # For date extraction in the 2nd plot of the Graphics section
```
</details>


<!-- DATA IMPORT -->
<details>
<summary><strong>Data</strong></summary>

This **data** is imported locally and the date field data type is set to a date type.

```{r data-import, class.source = 'fold-show'}
# Data import
file_path <- "../../Data/iosHealth/health_export_2026-01-12.csv"
health <- read_csv(file_path, col_types = cols(Date = col_date(format = "%m/%d/%y")))
```

<!-- DATA CLEANING -->
Lastly, the original **column names** were cleaned up.

<!-- Renmame Fields -->
```{r rename-fields, class.source = 'fold-show'}
# Rename fields
health <- health %>% 
  rename(
    Heart.Rate = `Heart Rate (average)`,
    Sleep.Sum = `Sleep (sum)`,
    Sleep.Core = `Sleep (Core) (sum)`,                
    Sleep.Deep = `Sleep (Deep) (sum)`,           
    Sleep.Asleep = `Sleep (Asleep) (sum)`,          
    Sleep.REM = `Sleep (REM) (sum)`,         
    Weight = `Weight (Most Recent) (most recent)`,
    Steps = `Steps (sum)`
  )
```

<!-- Add `Year` Column -->
  <!-- Extract year value from `Date` column. -->
```{r}
health <- mutate(health, Year = year(Date))
```
</details>

### Hide



<!--———————————————————————————————————————————————————————————
BACKGROUND
————————————————————————————————————————————————————————————-->
## Background {.tabset .tabset-pills .tabset-fade}

### Show

The [iPhone](https://www.apple.com/iphone/) in tandem with the [Apple Watch](https://www.apple.com/watch/) record numerous health metrics for the user/wearer made accessible in the [IOS Health App](https://www.apple.com/health/). Per Apple's promotions, users are intended to use these metrics to set goals and train more effectively. One intriguing metric to avid and casual health conscious users alike is daily step count. Measurements are more accurate for individuals who use an Apple Watch because a wearable devices tend to be on the user's person more consistently than they have their phone on them.

### Hide

</BR>


<!--———————————————————————————————————————————————————————————
ANALYSIS
————————————————————————————————————————————————————————————-->
## Analysis

#### Hypothesis

This analysis aims to model the actual daily total steps (*Total Steps*) of an individual according to their recorded daily average heart rate (*Avg Heart Rate*). Mathematically, 

<!-- TeX commands taken from:
https://byuistats.github.io/Statistics-Notebook/Analyses/Linear%20Regression/Examples/BodyWeightSLR.html# -->
$$
\underbrace{Y_i}_\text{Total Steps} = \beta_0 + \beta_1 \underbrace{X_i}_\text{Reported Avg Heart Rate} + \epsilon_i \quad \text{where} \ \epsilon_i \sim N(0, \sigma^2)
$$

The hypotheses to follow closely observe the slope of the regression model, $\beta_1$. If $\beta_1$ (slope) is 0, then we're confident that between *Total Steps* and reported *Avg Heart Rate* there is not a meaningful relationship.

$$
H_0: \beta_1 = 0 \\  
H_a: \beta_1 \neq 0 \\
\alpha = 0.1
$$



#### Graphics

The scatterplot below show the inter- and intra-year distribution of `Total Steps` by day from 2022-2026. There is a noticeable spike in the middle of 2024. The individual was employed to do outdoor work and averaged **≥ 20,000** steps daily for a few months from July-August.

<!-- Plot Setup: Base plot Variable: Daily Total Setps Over Time (5 yrs) -->
```{r}
# Base Plot
p_total_steps <-   ggplot(health, aes(x = Year, y = Steps, color = factor(Year))) +
    scale_color_brewer(palette = "Set1", name = "Year")+  # High contrast palette
    theme_bw() +
    theme(
      legend.position = "null"
    ) +
    labs(
      title = "Daily Total Steps Over Time (5 yrs)",
      subtitle = "Densest: 2023-2025; Spike: mid-to-late 2024"
    )

# Dataframe of counts
df_counts <- health %>% 
  group_by(Year) %>% 
  summarize(
    Points = n()
  )
```

<!-- Plot: Boxplot + Scatterplot: Daily Total Steps Over Time (5 yrs) -->
```{r}
p_total_steps +
  geom_boxplot() +
  geom_jitter(alpha = .1) +
  geom_label(data = df_counts, 
             aes(x = Year, y = Inf, label = Points),
             vjust = 1.5,
             show.legend = FALSE)+
  annotate("text", 
           x = 2022, 
           y = Inf, 
           label = "Points:", 
           vjust = 2,
           hjust = 1.5)
```

<!-- Plot: Distribution Curves -->
```{r}
# p_total_steps +
#   geom_density(aes(y=Steps, x = Year))
  # geom_jitter(alpha = .1) +
  # geom_label(data = df_counts, 
  #            aes(x = Year, y = Inf, label = Points),
  #            vjust = 1.5,
  #            show.legend = FALSE)+
  # annotate("text", 
  #          x = 2022, 
  #          y = Inf, 
  #          label = "Points:", 
  #          vjust = 2,
  #          hjust = 1.5)
```




<!--———————————————————————————————————————————————————————————
EXPLORATION 
————————————————————————————————————————————————————————————-->






<!--*********************************************************
NO LONGER USED
**********************************************************-->

```{r}
# WHAT: This code was from an original attempt to use exported IOS Health App data ingested as an XML and restructured as a workable dataframe. However, due to the size of the XML data file I turned on LFS (large file storage) to facilitate hosting my forked Statistics Notebook on GitHub. That removed the original XML file(s) and replaced them with hooks as LFS does. This however broke the connection I needed to use the data in RStudio for my analysis. I've since then changed gears and used a third party IOS to export the my health data directly as a CSV which will be used instead up above.
# 
# ```{r 0.1, echo=FALSE, message=FALSE, warning=FALSE}
# # ———————————————————————————————————————————————————————————————
# # DATA SOURCE (1): IOS Health App Data
# # ———————————————————————————————————————————————————————————————
# 
# 
# # STEP 1:
# ## Full file path: "/Users/andrewseaman/Desktop/2.STATS/Statistics-Notebook-master/Data/IOSHealth/apple_health_export/sleep-data.xml"
# 
# ## Load the XML file
# xml_file <- xmlParse("../../Data/IOSHealth/apple_Health_export/sleep-data.xml")# Replace with your file path
# root <- xmlRoot(xml_file)
# 
# 
# # STEP 2:
# ## Extract all 'Record' nodes
# records <- xmlToDataFrame(nodes = getNodeSet(root, "//Record"))
# 
# ## Filter only sleep data
# sleep_data <- records %>%
#   filter(type == "HKCategoryTypeIdentifierSleepAnalysis") %>%
#   select(type, startDate, endDate, value)
# 
# 
# # STEP 3:
# ## Convert to Date format
# sleep_data <- sleep_data %>%
#   mutate(
#     startDate = as_datetime(startDate),
#     endDate = as_datetime(endDate),
#     date = as_date(startDate)  # Extract only the date
#   ) %>%
#   filter(date >= as_date("2025-02-16") & date <= as_date("2025-02-25"))  # Filter target dates
# 
# 
# # STEP 4:
# # Rename sleep stages (Apple's values)
# sleep_data <- sleep_data %>%
#   mutate(sleep_stage = case_when(
#     value == "HKCategoryValueSleepAnalysisAwake" ~ "Awake",
#     value == "HKCategoryValueSleepAnalysisREM" ~ "REM",
#     value == "HKCategoryValueSleepAnalysisCore" ~ "Core",
#     value == "HKCategoryValueSleepAnalysisDeep" ~ "Deep",
#     TRUE ~ "Other"
#   ))
# 
# 
# # STEP 5:
# ## Summarize Total Sleep Durations Per Day
# sleep_summary <- sleep_data %>%
#   mutate(duration = as.numeric(difftime(endDate, startDate, units = "mins"))) %>%  # Compute duration
#   group_by(date, sleep_stage) %>%
#   summarise(total_minutes = sum(duration), .groups = "drop") %>%
#   pivot_wider(names_from = sleep_stage, values_from = total_minutes, values_fill = 0)  # Pivot for readability
# 
# 
# # STEP 6:
# ## View the data
# pander(sleep_summary)
# 
# ```
```




<!-- ************ -->
<!-- FILE END -->
<!-- ************ -->