---
title: "Logistic Regression"
subtitle: "Assesment Quiz"
author: "Andrew Seaman"
date: "2025-MM-DD"
output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

<!-- ------ -->
<!-- SETUP: -->
<!-- ------ -->

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load libraries
library(mosaic)
library(pander)
library(ResourceSelection)
library(tidyverse)
```

<!-- Setup End -->










<!-- ---------- -->
<!-- QUESTIONS: -->
<!-- ---------- -->

## 1.

Use an appropriate test to determine if the `birth weight` in ounces of a baby (wt) can be used to predict the probability that the `mother smoked` at all (smoke>0, or, if your Gestation data set has words in the "smoke" column, use smoke!="never") during or prior to the pregnancy. The graphic of the correct analysis is shown below.

```{r}
# a.
ges.glm <- glm( (smoke != "never") ~ wt, data = Gestation, family = binomial)

# b.
summary(ges.glm)

# c.
b <- coef(ges.glm)

# d.
plot( (smoke != "never") ~ wt, data = Gestation)
curve( exp(b[1] + b[2]*x)/(1 + exp(b[1] + b[2]*x)), add=TRUE)

```

> **Answer:**

\> ?Gestation

\> View(Gestation)

\> myglm <- glm(smoke>0 ~ wt, data=Gestation, family=binomial)

\> summary(myglm)

 </BR>

Call:
glm(formula = smoke > 0 ~ wt, family = binomial, data = Gestation)

Deviance Residuals:
    Min       1Q   Median       3Q      Max  
-1.8055  -1.2341   0.8984   1.0765   1.4810  

Coefficients:
             Estimate Std. Error z value Pr(>|z|)    
(Intercept)  2.416983   0.402280   6.008 1.88e-09 ***
wt              -0.018277   0.003311  -5.520 3.39e-08 ***

<!-- Q1. End -->










## 2.

I wasn't sure how to do this because I didn't have the original glm object. However, I knew you could calculate it by hand but wasn't fast enough to figure out that I had to use the match equation below.

```{r}
# a.
# COMMENT: I wasn't sure how to do this because I didn't have the original glm object. However, I knew you could calculate it by hand but wasn't fast enough to figure out that I had to use the match equation below.

# > I'm commenting this out because it won't compile otherwise.
# estimated_eb1 <- exp(ges.glm$coefficients["wt"])
# estimated_eb1
```

> **Answer:**

To compute a predicted value we can either do the calculation by hand:

$$
P(Y_i = 1|\, x_i) = \frac{e^{\beta_0 + \beta_1 x_i}}{1+e^{\beta_0 +
\beta_1 x_i}} = \pi_i
$$

`exp(-6.6035 + 0.3070*25) / (1 + exp(-6.6035 + 0.3070*25)) = 0.7448821`

Or we can use the predict function:

`predict(glmObject, data.frame(mpg=25), type="response")`

However, in this particular case, we don't have access to the "glmObject" that created the summary output, so we can only compute the answer by hand.
<!-- Q2. End -->





## 3.

I didn't quite get to this.

```{r}

```

> **Answer:**

To intepret the effect of mpg (the x-variable) on the odds, we compute exp(0.3070), which gives the value of eb1 . See the Overview page of Logistic Regression for an explanation.

Coefficients:
            Estimate Std. Error z value Pr(>|z|)   
(Intercept)  -6.6035     2.3514  -2.808  0.00498
mpg             0.3070     0.1148   2.673  0.00751

 

Since exp(0.3070) = 1.359341, we find that the odds of a success, i.e., that Yi =1 increases by 35.9341%, which rounds to 36% for every 1 unit increase in x.

Notices that the 35.9341 came from 1.359341. If you have 1.36 of what you originally had, then you increased by 0.36 or 36%.

<!-- Q3. End -->





<!-- Questions End -->




<!-- -------- -->
<!-- FILE END -->
<!-- -------- -->