---
title: "Class Activity - Good Peer Reviewing"
author: "Andrew Seaman"
date: "2025-01-13"
output: 
  html_document:
    theme: journal
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
# Load Libraries and View Data

library(mosaic)
library(tidyverse)
library(pander)
```

## Boxplot
Create a side-by-side boxplot where the y-axis shows the length of KidsFeet and the x-axis shows the gender of the child.
```{r}
boxplot(length ~ sex, 
  data=KidsFeet)
```

## Five-number Summary
Create a five-number summary table in support of the box plot.
```{r}
KidsFeet %>%
  group_by(sex) %>%
  summarise(            # Five-Number Summary:
    min=min(length),            # Min
    Q1=quantile(length, 0.25),  # Q1
    median=median(length),      # Median
    Q3=quantile(length, 0.75),  # Q3
    max=max(length)             # Max
  ) %>% 
  pander()
```

# Bar Plot
Create a bar plot that shows how many boys and how many girls are in the KidsFeet data set.
```{r}
barplot(table(KidsFeet$sex))
```

## Table of Counts
Create a supporting table of counts for the bar plot.
```{r}
table(KidsFeet$sex)
```

# Scatter plot
Create a scatter plot showing the Y-axis as length of `KidsFeet` and the X-axis as the width of `KidsFeet`.
```{r}
plot(length ~ width, data=KidsFeet)
```

## Correlation
Compute the correlation of the scatter plot.
```{r}
KidsFeet %>%
  summarise(Correlation = cor(width, length))
```

