---
title: "Day Care Fines"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---


```{r, warning=FALSE, message=FALSE}
library(mosaic)
library(DT)
library(pander)
library(car)
library(tidyverse)
library(lattice)

# To get the "late" dataset into your R Console, 
# you need to go to your file menu of R-Studio and select
# "Session, Set working directory, To source file location"
# Then click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 

## Wide data from original article:
late <- read.csv("../../../Data/late.csv", header=TRUE) 

## Change to "long" data for use in R:
Late <- pivot_longer(late, #Start with wide data set late
               cols = starts_with("Week"), #Find columns to gather from long data
               names_to = "Week", #New column name of Weeks in long data
               names_prefix = "Week", #Remove prefix of Week from Week1, Week2, ...
               names_transform = list(Week= ~as.numeric(.x)), #Make Week a numeric column of data
               values_to = "NumberofLateChildren") #Name of data column in long data

```



<!--If you want to give your critiquers some ideas about what you have questions on and would like help with, place those details here.-->



<!--End your comments to your critiquers before this line.-->

----


## Background {.tabset .tabset-fade}


### Study Details

<div style="color:#a8a8a8;">
This background is quoted directly from the article ["A Fine is a Price"](http://rady.ucsd.edu/faculty/directory/gneezy/pub/docs/fine.pdf). 
</div>

<div style="padding-left:30px; padding-right:30px;">

There are two types of day-care centers in Israel: private and public. A study was conducted in 10 private day-care centers in the city of Haifa from January to June 1998. All of these centers are located in the same part of town, and there is no important difference among them. During the day children are organized into groups according to age, from 1 to 4 years old. Each day-care center is allowed to hold a maximum of 35 children. In some exceptional cases a few additional children are allowed. The fee for each child is NIS 1,400 per month. (The NIS is the New Israeli Shekel.) At the time of the study, a U.S. dollar was worth approximately NIS 3.68, so the fee was about $380 at that time. 

The contract signed at the beginning of the year states that the day-care
center operates between 0730 and 1600. There is no mention of what happens if parents come late to pick up their children. In particular, before the beginning of the study, there was no fine for coming late. When parents did not come on time, one of the teachers had to wait with the children concerned. Teachers would rotate in this task, which is considered part of the job of a teacher, a fact that is clearly explained when a teacher is hired. Parents rarely came after 1630. 

A natural option [to fix the problem of parents showing up late] is to introduce a fine: every time a parent comes late, [they] will have to pay a fine. Will that reduce the number of parents who come late? If the fine is removed, will things revert back to the way they were originally?

The overall period of the study was 20 weeks. In the first 4 weeks we
simply recorded the number of parents who arrived late each week. At the
beginning of the fifth week, we introduced a fine in six of the 10 day-care centers, which had been selected randomly. The announcement of the fine was made with a note posted on the bulletin board of the day-care center. Parents tend to look at this board every day, since important announcements are posted there. The announcement specified that the fine would be NIS 10 for a delay of 10 minutes or more. The fine was per child; thus, if parents had two children in the center and they came late, they had to pay NIS 20. Payment was made to the principal of the day-care center at the end of the month. Since monthly payments are made to the owner during the year, the fines were added to those amounts. The money was paid to the owner, rather then to the teacher who was staying late (and did not get any additional money). The teachers were informed of the fine but not of the study. Registering the names of parents who came late was a common practice in any case.

At the beginning of the seventeenth (17th) week, the fine was removed with no explanation. Notice of the cancellation was posted on the board. If parents asked why the fines were removed, the principals were instructed to reply that the fine had been a trial for a limited time and that the results of this trial were now being evaluated.

A comparison with other fines in Israel may give an idea of the size of
the penalty that was introduced. A fine of NIS 10 is relatively small but not insignificant. In comparison, the fine for illegal parking is NIS 75; the fine for driving through a red light is NIS 1,000 plus penalties; the fine for not collecting the droppings of a dog is NIS 360. For many of these violations, however, detection and enforcement are low or, as in the case of dog dirt, nonexistent in practice. A baby-sitter earns between NIS 15 and NIS 20 per hour. The average gross salary per month in Israel at the time of the study was NIS 5,595.

</div>

</BR>

**Make note of the following:**

*How many Day Care Centers were included in the study?*

* 10 (type: private)

*How many Centers were included in the treatment group? The Control Group?*

* Treatment Group: 6
* Control Group: 4

*During which weeks did the researchers simply observe the Day Care Centers?*

* Weeks: 1-4

*During which weeks was the fine applied to the treatment group?*

* Weeks: 5-16

*During which weeks was the fine removed from the treatment group?*

* Weeks: 17-20

*What was the main response variable of interest that the researchers recorded each week on the Day Care Centers?*

* Parents showing up late to pick up their children.

### The Data (Wide)

The `late` Day Care Center data is shown here in the "wide data format" that was provided by the authors of the original research article. But this format is not terribly useful for an ANOVA analysis.

```{r, out.width=300}
#Show the full width of the "Wide" version of the late data:
pander(late, split.tables = Inf)
```


### The Data (Long)

The `Late` Day Care Center data is shown here in the "long data format", a more usable format for analysis.

```{r}
pander(Late)
```


### Add Factor

**Instructions:**

Use the Late data set to add a new column to the data set that translates Weeks (a numeric column) into a column that is a factor called "Period". This factor Period should have three levels: "Before the Fine", "Fine Applied", and "Fine Removed".

Hint: Look up the mutate( ) function on the R Commands page of the Statistics-Notebook and study the second Example Code that creates a new column called "season" from the numeric "birthmonth" column in the KidsFeet dataset.

```{r}
# Add new column. Translate 'Week' into a factor.
Late <- Late %>% 
  mutate(
    Period = factor(
      case_when(
        Week >= 1 & Week <= 4   ~ "Before the Fine",
        Week >= 5 & Week <= 16  ~ "Fine Applied",
        Week >= 17 & Week <= 20 ~ "Fine Removed"
      ),
      levels = c("Before the Fine", "Fine Applied", "Fine Removed")
    )
  )
```

**Test:**

```{r}
table(Late$Period) %>% pander()
```

 
##  


----



## Analysis {.tabset .tabset-fade}

### Hypotheses

<!-- State the hypotheses for this analysis in this section. -->

#### Set 1:

Hypotheses about Treatment where $\mu$ represents the mean number of late children picked up from Day Care Centers during the week.

$$
H_0: \mu_\text{Control Group} = \mu_\text{Fine Group}
\\
H_a: \text{The average number of late children differs for one Group}
$$

#### Set 2:

Hypotheses about Period where $\mu$ represents the mean number of late children picked up from Day Care Centers during the week.

$$
H_0: \mu_\text{Before the Fine} = \mu_\text{Fine Applied} = \mu_\text{Fine Removed}
\\
H_a: \text{The average number of late children differs for at least one Period}
$$

#### Set 3:

Hypotheses about the interaction of Treatment and Period.

$$
H_0: \text{The effect of the Group on the average number of late children picked up does not depend on the Period of the study.}
\\
H_a: \text{The effect of the Group on the average number of late children picked up changes during at least one Period of the study.}
$$



### Two-way ANOVA 

<!-- Show the two-way ANOVA summary output in this section. -->

```{r}
# Two-Way ANOVA: 

  # Notation: myaov <- aov(Y ~ X1+X2+X1:X2, data=YourDataSet)

myaov <- aov(NumberofLateChildren ~ Treatment + Period + Treatment:Period, data = Late)

results <- summary(myaov)
```
**Raw Output:**
```{r}
results
```

**Clean Output:**
```{r}
pander(results)
```


*Which p-values were significant?*

* All of them (Treatment, Period, Treatment:Period)


### Graphics & Numerical Summaries {.tabset .tabset-fade}

<!-- Show the three graphics of the two-way ANOVA in this section and provide a supporting table of numerical summaries for each graphic. -->

#### Graphic 1

```{r}
# Graphic 1: First, create an xyplot(...) of the Treatment factor. Creating a supporting table of means for this plot.

  # Notation: plot(y ~ x, data=NameOfYourData)
    # y is the quantitative response variable, i.e., “numeric vector.”
    # x is the quantitative explanatory variable, i.e., “numeric vector.”
    # NameOfYourData is the name of the dataset containing y and x.

xyplot(NumberofLateChildren ~ as.factor(Treatment),
       data = Late,
       type = c("p", "a"),  # points and averages
       main = "Late Children by Treatment",
       xlab = "Treatment Group",
       ylab = "Number of Late Children")
```

```{r}
# Numerical Summary 1: 
numSummary1 <- Late %>% 
  group_by(Treatment) %>% 
  summarise('Mean Late Children' = mean(NumberofLateChildren, na.rm = TRUE)) %>% 
  ungroup()

pander(numSummary1)
```

#### Graphic 2

```{r}
# Graphic 2: Second, create an xyplot(...) of the Period factor and include a supporting numerical summary table of means.

xyplot(NumberofLateChildren ~ as.factor(Period),
       data = Late,
       type = c("p", "a"),  # points and averages
       main = "Late Children by Period",
       xlab = "Period",
       ylab = "Number of Late Children")
```

```{r}
# Numerical Summary 2: 

numSummary2 <- Late %>% 
  group_by(Period) %>% 
  summarise('Mean Late Children' = mean(NumberofLateChildren, na.rm = TRUE)) %>% 
  ungroup()

pander(numSummary2)

```

#### Graphic 3

```{r}
# Graphic 3: Third, and most importantly, create an xyplot(...) that shows the interaction of the Treatment and Period variables. It is best practice to place the factor with the most levels as the x-axis of the graph (Period in this case) and the factor with the fewest variables as the legend (Treatment in this case). Then, create a table of numerical summaries showing the means for all six conditions presented in the plot.

xyplot(NumberofLateChildren ~ Period,
       data = Late,
       group = Treatment,   # sets legend groups
       type = c("p", "a"),  # points and averages
       auto.key = TRUE,     # automatically generates legend
       main = "Late Children by Treatment and Period",
       xlab = "Period",
       ylab = "Number of Late Children")
```

```{r}
# Numerical Summary 3: 

numSummary3 <- Late %>% 
  group_by(Treatment, Period) %>% 
  summarise(
    `Mean Late Children` = round(mean(NumberofLateChildren, na.rm = TRUE), 2),
    .groups = "drop"
  )

pander(numSummary3)
```

###

### Conclusions

<!-- Interpret the results of the study. Use specific details in your writing like mentioning the actual amount of the fine and the relative cost to babysitting. Mention the average number of late children before the fine, during the fine, and after the fine was removed. -->

The two-way ANOVA results show statistically significant effects for Treatment (p < 0.001), Period (p < 0.001), and the Treatment × Period interaction (p < 0.001). This means:
	1.	The number of late pickups differed significantly between the Control and Fine groups overall.
	2.	The number of late pickups also differed significantly between the three Periods (Before the Fine, Fine Applied, Fine Removed).
	3.	Most importantly, the effect of Treatment depended on the Period — the difference between Control and Fine groups changed over time.

From the numerical summaries:
	•	Before the fine was applied, both groups had similar late pickups (Control ≈ 10, Fine ≈ 8).
	•	Once the fine of NIS 10 (~$2.72 USD) per child for delays ≥10 minutes was introduced, late pickups increased sharply in the Fine group to ≈ 16.4, while the Control group stayed roughly the same (≈ 9.2).
	•	After the fine was removed, the Fine group’s late pickups rose further to ≈ 18.7, while the Control group dropped slightly to ≈ 8.3.

The interaction plot reinforces this pattern: the Fine group’s trend line rises steeply after the fine is introduced, while the Control group’s line stays flat or slightly decreases.

These results suggest that instead of deterring late pickups, the fine may have normalized or even incentivized lateness in the Fine group, possibly because parents reframed lateness as a service they could “purchase” cheaply — especially compared to babysitting costs (NIS 15–20/hour, ≈$4.08–$5.44 USD). The removal of the fine did not reverse this effect; lateness remained higher than before the fine was introduced.

The fine not only **failed** to reduce late pickups, but it significantly increased them, and this effect persisted after its removal.




### Diagnostic Plots

<!-- Provide some discussion about how appropriate the ANOVA analysis was for this data. Include the residuals vs. fitted values plot and the Q-Q Plot of residuals. -->

How do you check if the ANOVA assumptions are satisfied?

* You must use a QQ-plot to assess the normality of the data.

What do you say about your conclusions if the assumptions aren't satisfied?

* We cannot accept these conclusions because the results will be heavily swayed by non-normal data in any of the groups and non-equal variance between the groups. For comparison sake, the groups must be fairly similar in how they each distribute themselves and each distribution must be normal.





