---
title: "Heaven On Earth Massage"
subtitle: "BYU-Idaho Consulting Project (MATH325)"
author: "Andrew Seaman"
date: "03/16/2026"
output:
  html_document:
    theme: flatly
    code_folding: hide
    toc: true
    toc_float: true
    toc-depth: 6
---

<!--******************************************************
LAST UPDATED: 03/30/2026
BY: Andrew Seaman
*******************************************************-->

<!-------------------------------------------------------
.                          SETUP                    
-------------------------------------------------------->

```{r imports, message=FALSE, warning=FALSE}
# Import libraries
pacman::p_load( ggplot2, ggridges, readr, mosaic, directlabels, DT, lubridate, pander, plotly, tidyverse)

# Import dataset
mssg <- read_csv("../Data/massage.csv", col_types = "dcccdcdcddddd") %>% 
  rename(
    week = "Week",
    date = "Date",
    staff = "Staff",
    client = "Client",
    description = "Description",
    length_min = "Length (min)", 
    base_amt = "Base Amt ($)", 
    discount_pct = "Discount (%)", 
    discount_amt = "Discount ($)", 
    paid_amt = "Paid Amt ($)", 
    tip_amt = "Tip Amt ($)", 
    total_amt = "Total Amt ($)"
  ) %>% 
  # Add calculated columns & Clean data types
  mutate(
    length_hr = length_min/60,
    discount_pct = readr::parse_number(discount_pct) / 100,
    day = substr(date,1,3),
    date = lubridate::mdy(substr(date, 5, nchar(date)))
  )
```

## {.tabset .tabset-pills .tabset-fade}

### Hide

<!-------------------------------------------------------
.                          PLANNING                    
-------------------------------------------------------->

<!-- Analysis Ideas -->
```{r, echo=FALSE, include=FALSE}
# ============================================================
#                           INTRO
# ============================================================
# From the 'Class Activity - Idea Approval for the Week 11 "Consulting Opportunity or Research Project"' I've complied the following ordered lists for clarify and reference. 

# However, I believe I've already found my analysis idea and client.

# ************************************************************


# ————————————————————————————————————————————————————————————
#                 PAST CLIENT-FOCUSED ANALYSES
# ————————————————————————————————————————————————————————————
# 	1.	Identified truck drivers with the best and worst live-beehive delivery performance by analyzing survival rates, delivery times and distances (for a family beekeeping business).
# 	2.	Assessed weather’s effect on chicken egg-laying rates and recommended climate-control improvements (for a family chicken farm).
# 	3.	Contributed data analysis to a grant renewal proposal by quantifying improvements in student mental-happiness rates over three years (for a high school).
# 	4.	Saved a construction company thousands by modeling the optimal number of employees per job (for a family construction business).
# 	5.	Pinpointed which course assignments most critically impacted student success to guide TA support efforts (as a teaching assistant).
# 	6.	Measured how an outdoor stand-up sign influenced walk-in customer traffic (for a local retail business).
# 	7.	Ran a profit analysis to identify top-revenue items and recommend discontinuations or production increases (for a local retailer).
# 	8.	Performed a cost-benefit study comparing tanning-bed profits versus manicure rooms—leading to a $250/month profit boost after removing the tanning bed (for a neighborhood salon).

# ————————————————————————————————————————————————————————————
#                 GRADUATE-SCHOOL RESEARCH PROJECTS
# ————————————————————————————————————————————————————————————
# 	1.	Mouse example logistic-regression analysis (see textbook walkthrough: https://byuistats.github.io/Statistics-Notebook/Analyses/Logistic%20Regression/Examples/mouseLogisticReg.html).
# 	2.	NHANES-based study on how depression and poverty affect diabetes diagnosis probability.
# 	3.	Evaluation of ABiC glaucoma surgery’s impact on intraocular pressure and medication needs (for publication by an Idaho Falls optometrist).

# ————————————————————————————————————————————————————————————
#                 SELF-DRIVEN ANALYSES
# ————————————————————————————————————————————————————————————
# 	1.	Determining which factors most influence acceptance into data-science graduate programs.
# 	2.	Tracking psychology graduates’ post-degree paths (jobs, further study) and key characteristics linked to employment success.
# 	3.	Monthly budgeting: spending breakdown and potential cut-back areas.
# 	4.	Comparative value analysis of baby strollers for cost-effectiveness.
```

<!-- Chosen Analysis -->

</BR>

---


### Data Exploration {.tabset .tabset-pills .tabset-fade}

**(EXPLORE THE DATA TO FIND SOMETHING TO ANALYZE)**

#### Uniqe values

Show all unique values by column.

```{r}
unique_column_value_lists <- mssg %>%
  summarise(across(everything(), ~ paste(unique(.), collapse = ", "))) %>% 
  pander()

unique_column_value_lists
```

</BR>

Show a table of counts of unique values by column.

```{r}
# table_of_counts_all_columns <- mssg %>%
#   tidyr::pivot_longer(everything(), names_to = "column", values_to = "value") %>%
#   dplyr::count(column, value)
# 
# table_of_counts_all_columns
```

</BR>

#### Plots

Pepare a table of counts and convert it into a dataframe. (This is most likley a very inefficient way to wrangle this data).

```{r}
tbl_counts <- table(mssg$staff, mssg$week)
df_counts <- as.data.frame(tbl_counts) %>% 
  rename(
    staff = Var1,
    week = Var2,
    count = Freq
  ) %>%
  mutate(
    staff = as.character(staff),
    weeks = as.numeric(week)
  )
```

**Plot 1:** Horizontal line plot of massages given per week from 2025-2026.
```{r}
# 1 col, 5 rows
ggplot(df_counts, aes(x=week, y=count, color=staff, group=1))+
  geom_line()+
  facet_wrap(~staff, ncol = 1)
```

**Plot 2:** Vertical line plot of massages given per week from 2025-2026.
```{r}
# 5 cols, 1 row
ggplot(df_counts, aes(x=week, y=count, color=staff, group=1))+
  geom_line()+
  facet_wrap(~staff, ncol = 5)+
  coord_flip()
```

**Plot 3:** Ridge plot of massages given per week from 2025-2026.
```{r}
# Ridge Plot
# ggplot(df_counts, aes(x=week, y=staff, fill ))+
#   geom_density_ridges()


ggplot(mssg, aes(x=week, y=staff, fill = staff))+
  geom_density_ridges()+
  theme_ridges()+
  geom_text(aes(label= count(week)))
```

**Plot 4:** Clean ridge plot marking the mean and median of massages given in 2025-2026.
```{r, fig.height=8, fig.width = 10, fig.align="center"}
mssg_clean <- mssg %>%
  filter(!is.na(staff), !is.na(week))

week_counts <- mssg_clean %>%
  count(staff, week)

max_week <- max(mssg$week)
fill_colors <- palette.colors(n = 5, palette = palette.pals()[8]) 

# FAVORITES: ["Tableau 10", "", "", "", ""]
#  [1] "R3"              "R4"              "ggplot2"         "Okabe-Ito"       "Accent"          "Dark 2"         
#  [7] "Paired"          "Pastel 1"        "Pastel 2"        "Set 1"           "Set 2"           "Set 3"          
# [13] "Tableau 10"      "Classic Tableau" "Polychrome 36"   "Alphabet"  

ggplot(mssg_clean, aes(x = week, y = staff, fill = staff)) +
  labs(
    title = "Seasonal Highs and Lows Are Apparent For All Massage Therapists",
    subtitle = "Stats: red circle for mean; green point for median",
    x = "Week",
    y = "Massage Therapist"
  ) +
  
  # Geom - Section:
  geom_vline(xintercept = c(12, 24, 36, 48), linetype = "dotted", show.legend = FALSE) +
  geom_density_ridges(
    scale = .9,
    alpha = 0.7,
    color = "black",
    rel_min_height = .001
  ) +
  geom_point(
    # pch = 20, 
    shape = 21,
    size = 7, 
    alpha = .4,
    stat = "summary",
    fun = mean,
    color = "red"
  ) +
  geom_point(
    pch = 20, 
    size = 7, 
    alpha = .4,
    stat = "summary",
    fun = median,
    color = "green"
  ) +
  geom_label(
    aes(label = staff), 
    stat = "summary", 
    fun = median, 
    vjust = -.8, 
    hjust = c(4.7, 7.3, 2.9, 4, 4.6)) +
  
  annotate("text", x = c(6,18,30,42), y = .5, label = c("Q1", "Q2", "Q3", "Q4"))+
  
  # Scale - Section:
  scale_x_continuous(limits = c(1,53), breaks = seq(1,53, by = 1)) +
  scale_fill_manual(values = fill_colors) +
  
  # Theme - Section:
  theme_ridges() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(size = 6),
    axis.text.y = element_blank(),
    # NOTE: `theme_ridges()` may be interfering with the following line of code.
    # axis.title = element_text(hjust = 0.5),
    axis.title.x = element_text(hjust = 0.5),
    axis.title.y = element_text(hjust = 0.5)
  )

# NOTE: values to pass to `stat = ...`
# 
# "identity"    # use raw y values
# "count"       # count rows (default for geom_bar)
# "bin"         # histograms
# "density"     # density curves
# "smooth"      # smoothing models
# "summary"     # summary stats (mean, etc.)
# "summary_bin" # binned summaries
# "boxplot"     # boxplot stats
# "ydensity"    # violin plots
# "ecdf"        # empirical CDF
# "qq"          # QQ plot
# "qq_line"

# NOTE: values to pass to `fun = ...`
# 
# fun = mean
# fun = median
# fun = sum
# fun = min
# fun = max
# fun = length      # count
# fun = sd
# fun = var
```

**Plot 5:** Vertical bar plot of massages given per week from 2025-2026.
```{r, fig.height=8, fig.width = 10, fig.align="center"}
# Bar chart - Get a clearer picture of what the ridgeplot is showing
ggplot(mssg_clean, aes(x = week, fill = staff)) +
  labs(
    title = "Massive gap from week 4 to week 26 YoY (2025-2026).",
    subtitle = "Note: Emma and Luke were terminated",
    x = "Week",
    y = "Count",
    caption = "Figure #. shows how many clients each massage therapist saw and when by week number throughout the year."
  ) +
  geom_bar(position = "dodge") +
  scale_x_continuous(limits = c(1,53), breaks = seq(1,53, by = 1)) +
  scale_fill_manual(values = fill_colors)+
  facet_wrap(~staff, ncol = 5) +
  theme(
    legend.position = "none",
    axis.text = element_text(size = 6)
  ) +
  coord_flip()
```

**Plot 6:** Line plot with slight line curve for massages given per week from 2025-2026.
```{r}
df <- mssg_clean %>% 
  group_by(week, staff) %>% 
  summarise(count = n())

ggplot(df, aes(x=week, y=count, color=staff))+
  # geom_line()
  geom_smooth(se = FALSE, method = "glm", method.args = list(family = "poisson"))+
  geom_jitter(width = 0, alpha = .5)
  
```


</BR>

*(EXPLANATION)*

```{r}
# (Description)

```

</BR>

*(EXPLANATION)*

```{r}
# (Description)

```

</BR>

*(EXPLANATION)*

```{r}
# (Description)

```

---






## Background 

(BACKGROUND TEXT)

[Heaven On Earth](https://www.massagebook.com/therapists/sleepyridge?src=external) is a growing massage therapy spa located in Orem, Utah. Services begin with a basic massage of 60 minutes and can increase in time and intensity as needed. Add-on services are also provided upon request. An important focus emphasized by each masseuse is to come in with a specific intention. That may mean one focuses on overall wellness, stress relief, pain and tension release, etc.

</BR>

---



## Question and Hypothesis

When it comes time to pay for the services provided what motivates individuals to add a tip? Could the most important lever be simply which masseuse provided the services?

*(STATE HYPOTHESIS)...ANOVA?*

$$
H_0:
\\
H_a:
\\
\alpha = 0.10 \ \text{(90% CL)}
$$

</BR>



## Data

```{r libs, message=FALSE, warning=FALSE}
# View Data
datatable(
  mssg,
  options = list(
    lengthMenu = list(c(5, 10, -1), c("5", "10", "All"))
  )
)
```

---

</BR>



## Fitting the Model

*(STATISTICAL TESTING)*

```{r}
# (Statistical Test Name...tbd)

```

*(INTERPRETATION)*

</BR>



> ### Diagnostic Plots

*(TEST FOR MODEL ASSUMPTIONS)*

```{r}
# Diagnostic Plots

```

*(INTERPRETATION)*

</BR>

---



## Graphics

**(GRAPHICS, NUMERICAL SUMMARIES, AND INTERPRETATION)**

<!-- G1 -->
*(INTERPRETATION)*

```{r}
# Graphic 1

```

<!-- NS1 -->
*(INTERPRETATION)*

```{r}
# Numerical Summary 1

```

</BR>

<!-- G2 -->
*(INTERPRETATION)*

```{r}
# Graphic 2

```

<!-- NS2 -->
*(INTERPRETATION)*

```{r}
# Numerical Summary 2

```

</BR>

<!-- G3 -->
*(INTERPRETATION)*

```{r}
# Graphic 3

```

<!-- NS3 -->
*(INTERPRETATION)*

```{r}
# Numerical Summary 3

```

</BR>

---





## Conclusion

---

*(SUMMARY)*






</BR></BR></BR></BR></BR>

<!-- ======== -->
<!-- FILE END -->
<!-- ======== -->