---
title: "One-Way ANOVA"
subtitle: "Class Activity"
author: "Andrew Seaman"
date: "2025-Feb-10"
output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
    code_folding: hide
---

<!-- ------ -->
<!-- SETUP: -->
<!-- ------ -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load libraries
library(mosaic)
library(DT)
library(pander)
library(tidyverse)
```

```{r}
# Import dataset
kf <- KidsFeet
```

<!-- Setup End -->





<!-- ---------- -->
<!-- QUESTIONS: -->
<!-- ---------- -->

## Questions

Check off the items below as you complete them during today's class:

- [x] I have located the ANOVA content of my Statistics Notebook under the Making Inference menu. 
- [x] I can explain when to use each of the t Tests, each of the Wilcoxon Tests, and ANOVA. 
- [x] I can explain the words "factor" and "level" using the airquality dataset as an example. 
- [x] Using the word "factor" in my explanation, I can explain the difference between a one-way and a two-way ANOVA. 
- [x] I have reviewed the chickwts (One-way) and warpbreaks (Two-way) example analyses in R and understand how to complete all of the R Code in those files.

<!-- Questions End -->





<!-- ---------- -->
<!-- NOTES: -->
<!-- ---------- -->

## Factor

**Definition:** A factor is defined as a qualitative variable containing at least two categories. The categories of the factor are referred to as the “levels” of the factor.

By using `as.factor(x)` on the categorical variable we can treat it as data type factor.

View the data types with either `str()`.`glimpse()`.

```{r}
# a. `glimpse()`
glimpse(kf) %>% pander()

```



## ANOVA

**Notation:**

You have to run the analysis (`aov()`) and assign it to a variable first (e.g. `myaov`) and then view the summary (`summary(myaov)`).

**One-Way ANOVA:** `myaov <- aov(Y ~ X, data=YourDataSet)`

**Two-Way ANOVA:** `myaov <- aov(Y ~ X1+X2+X1:X2, data=YourDataSet)`

**Summary:** `summary(myaov)`


<!-- Notes End -->





<!-- -------- -->
<!-- FILE END -->
<!-- -------- -->