---
title: "Permutation Testing (Part 2)"
subtitle: "Class Activity"
author: "Andrew Seaman"
date: "2026-04-06"
output: 
  html_document:
    # highlight-style: darkly 
      # Highlight Styles:   cosmo (clean, modern), flatly, litera, minty, lux, sandstone, journal,
      #                     darkly (dark mode), cyborg (very dark hacker vibes), github
    theme: simplex  
      # Themes:             default, bootstrap, cerulean, cosmo, darkly, flatly, journal, lumen,
      #                     paper, readable, sandstone, simplex, spacelab, united, yeti
    toc: true
    # toc_float: true
    code_folding: show
---

<!-- IMPORTS -->
```{r, message=FALSE, warning=FALSE, include=FALSE}
# Libraries
pacman::p_load(mosaic, DT, pander, tidyverse)
```

---

<!-- ————————————————————————————————————————————————————————————————— -->


## Approaches {.tabset .tabset-fade .tabset-pills}

### T-Test

**Step 1**
<!--
# SOURCE: Stats Textbook & Class Activity
-->
```{r}
i <- "> 1. First run the initial test and gain the test statistic:"
# myTest <- t.test(extra ~ group, data = sleep, mu = 0)
# observedTestStat <- myTest$statistic

# > Data: Filter and keep only 4 & 8 cylinders
df <- mtcars %>%
  filter(cyl %in% c(4, 8)) %>% 
  rename(
    cylinder = cyl,
    weight = wt
  )

# > Run an initial t-test
myTest <- t.test(weight ~ cylinder, data = df)
myTest %>% pander()

# > Get the test statistic
observedTestStat <- myTest$statistic
observedTestStat
```

**Step 2**
```{r}
i <- "> 2. Now we run the permutations to create a distribution of test statistics"

# N <- 2000
# permutedTestStats <- rep(NA, N)
# for (i in 1:N){
#   permutedTest <- t.test(sample(extra) ~ group, data = sleep, mu = 0)
#   permutedTestStats[i] <- permutedTest$statistic
# }

N <- 2000
permutedTestStats <- rep(NA, N)
for (i in 1:N) {
  permutedTest <- t.test(sample(weight) ~ cylinder, data = df, mu = 0)
  permutedTestStats[i] <- permutedTest$statistic
}
```

##### {.tabset .tabset-fade .tabset-pills}

###### Hide


###### Show Output

2000 permutations simulating the data we need to be able to see if the number of cylinders is independent from the weight of the vehicle or not.
```{r}
permutedTestStats
```

####

**Step 3**
```{r}
i <- "> 3. Now we show a histogram of that distribution"

# hist(permutedTestStats, col = "skyblue")
# abline(v = observedTestStat, col = "red", lwd = 3)

hist(permutedTestStats, col = "skyblue",
     xlim = range(c(permutedTestStats, observedTestStat)))
abline(v= observedTestStat, col = "red", lwd = 3)
```

**Step 4**
```{r}
i <- "> 4.1/3. Greater-Than p-value: Not the correct one in this case."
# sum(permutedTestStats >= observedTestStat)/N
sum(permutedTestStats >= observedTestStat)/N


i <- "> 4.2/3. Less-Than p-value: Not the correct one for this data"
# sum(permutedTestStats <= observedTestStat)/N
sum(permutedTestStats <= observedTestStat)/N


i <- "> 4.3/3. Two-Sided p-value: This is the one we want based on our alternative hypothesis."
# 2*sum(permutedTestStats <= observedTestStat)/N
2*sum(permutedTestStats <= observedTestStat)/N
```

---

<!-- ————————————————————————————————————————————————————————————————— -->





### One-Way ANOVA

Perform a One-way ANOVA using a permutation test.  Use the diamonds dataset and test whether the average price of the diamonds depends on the clarity.  Use the following psuedo code as a starting spot.  (Hint: because the diamonds dataset has 53,000 rows we start with a small N value)

<!-- One-Way ANOVA -->
```{r}
# SOURCE: Class Activity

# > Data
df <- diamonds

#Step 1
myTest <- aov(price ~ clarity, data = df)
observedTestStat <- summary(myTest)[[1]]$`F value`[1]


#Step 2
N <- 2000 #53000 - I've used: 500, 2000, 10000, and 53000 (which ran for about a minutes)
permutedTestStats <- rep(NA, N)
for (i in  1:N){
  permutedTest <- aov(sample(price) ~ clarity, data = df)
  permutedTestStats[i] <- summary(permutedTest)[[1]]$`F value`[1]
}

hist(permutedTestStats, col = "skyblue")
abline(v=observedTestStat, col = "red", lwd = 3)
  
#Step 3
sum(permutedTestStats >= observedTestStat)/N
sum(permutedTestStats <= observedTestStat)/N
```

**Observed Test Statistic:**
```{r}
observedTestStat
```


**TEMP TEST CHUNK! DELETE!**
```{r}
#     TEMP! DELETE THIS SOON! <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<----------------------------- *******
#
#     

ggplot(diamonds, aes(x = price)) +
  # geom_histogram() +                                  #1
  
  # geom_jitter(aes(y = clarity))                       #2
  
  # geom_density(aes(fill = clarity, alpha = .1))       #3
  
  geom_boxplot(aes(y = clarity)) +                    #4
  coord_flip()
  
  # geom_violin(aes(y = clarity)) +                       #5
  # coord_flip()
```


---

<!-- ————————————————————————————————————————————————————————————————— -->





### Logistic Regression

Perform a Logistic Regression Test using a permutation test.  Use the SAT dataset (mosaic library) and test whether the likelihood of scoring above 1000 (sat > 1000) on the SAT depends on the expenditure per pupil (expend).  Use the following psuedo code as a starting spot.

<!-- Logistic Regression -->
```{r}
# SOURCE: Class Activity

# > Data: filter and keep only two levels of 
df <- SAT %>% 
  mutate(
    sat_dummy = ifelse(sat > 1000, 1, 0)
  )

# #Step 1
myTest <- glm(sat_dummy ~ expend, data = df, family = binomial)
observedTestStat <- summary(myTest)[[12]][2,3]

# #Step 2
N <- 100
permutedTestStats <- rep(NA, N)
for (i in  1:N){
  permuted_df <- df %>% 
    mutate(sat_dummy = sample(sat_dummy))
  
  permutedTest <- glm(sat_dummy ~ expend, data = permuted_df, family = binomial)
  permutedTestStats[i] <- summary(permutedTest)[[12]][2,3]
}

hist(permutedTestStats, col = "skyblue")
abline(v=observedTestStat, col = "red", lwd = 3)

#Step 3
sum(permutedTestStats >= observedTestStat)/N
sum(permutedTestStats <= observedTestStat)/N
```

**Observed Test Statistic:**
```{r}
observedTestStat
```

---

<!-- ————————————————————————————————————————————————————————————————— -->





<!-- -------- -->
<!-- FILE END -->
<!-- -------- -->