---
title: "Randomization Testing"
subtitle: "Skills Quiz"
author: "Andrew Seaman"
date: "2026-04-09"
output: 
  html_document:
    theme: simplex
    code_folding: hide
    toc: true
    toc_depth: 6
    collapsed: true
    smooth_scroll: true
---

<!-- IMPORTS -->
```{r, message=FALSE, warning=FALSE, include=FALSE}
#libraries
pacman::p_load(mosaic, car, DT, pander, tidyverse)
```

<!-- START: ALL -->

<!-- START: Main Tabset -->
# Pages {.tabset .tabset-pills .tabset-fade}





<!-- START: Summary -->
## Summary


### Introduction
[Introduction text]

<BR />


### Questions

##### Q1
A quick check before proceeding.

##### Q2
Match the null hypothesis to the test.

##### Q3


##### Q4
Freebie.

##### Q5


##### Q6


<!-- END: Summary -->





<!-- START: Q1 -->
## Q1
**Question:** Have you completed both class activities for this week? 

**Answer:** Yes, I have.

<!-- END: Q1 -->





<!-- START: Q2 -->
## Q2 

Match the following Tests to their Null Hypothesis.

1. **Independent Samples t Test**
$$H_0: \mu1=\mu2$$
 

2. **Wilcoxon Rank Sum Test**
$$H_0: \text{difference in medians} = 0$$

3. **Chi-squared Test**
$$H_0: \text{the two variables are not associated}$$

4. **ANOVA**
$$H_0:\mu1=\mu2=\mu3=\mu4$$
 
5. **Wilcoxon Signed-Rank Test**
$$H_0: \text{median of the differences} = 0$$

6. **Paired Samples t Test**
$$H_0: \mu d=0$$

<!-- END: Q2 -->





<!-- START: Q3 -->
## Q3

##### 3.1
Run the following codes in R.

Here we have the first five rows of the random data we just generated, sampled, and assembled into a workable dataset.
```{r}
# 3.1

#random data
set.seed(1140411) #allows us each to get the same set of random values

#samples
sample1 <- rnorm(30, 69, 2.5) #get a random sample of n=30 normally distributed values with mu=69 and sigma=2.5

sample2 <- rnorm(30, 69, 2.5) #get another random sample of n=30 normally distributed values with mu=69 and sigma=2.5

#data frame
theData <- data.frame(values = c(sample1,sample2), group = rep(c(1,2), each=30)) #load the random samples into a data set

#head
head(theData, 5) %>% pander()
```

##### 3.2
Boxplot
```{r}
# 3.2

#plot
boxplot(values ~ group, data = theData)
```

##### 3.3
We will run an Independent Samples T-test.
```{r}
# 3.3

#independent samples t-test
myTest <- t.test(values ~ group, data = theData, mu = 0)
myTest %>% pander()
```

##### 3.4
Observed Test Statistic.
```{r}
# 3.4

observedTestStat <- myTest$statistic
observedTestStat %>% pander()
```

##### 3.5
Permutation test.
```{r}
# 3.5

#permutation test
N <- 2000      
permutedTestStats <-  rep(NA, N)
for  (i in 1:N ) {
   permutedTest <- t.test(sample(values) ~ group, data = theData, my = 0)

   permutedTestStats[i]  <-  permutedTest$statistic
  
}
hist(permutedTestStats)
abline(v=observedTestStat)
sum(permutedTestStats >= observedTestStat)/N
sum(permutedTestStats <= observedTestStat)/N
```

<!-- END: Q3 -->





<!-- START: Q4 -->
## Q4
This question has been deleted due to an important error in the question.

If you would like credit for this question, click "Yes" below:

<!-- END: Q4 -->





<!-- START: Q5 -->
## Q5 {.active}

##### 5.1
```{r}
# 5.1

#table
table(SaratogaHouses$fuel)
```

##### 5.2
```{r}
# 5.2

#kruskal-wallis test
kruskalTest <- kruskal.test(price ~ fuel, data = SaratogaHouses)
kruskalTest %>% pander()
```

##### 5.3
```{r}
# 5.3

#plot
boxplot(price ~ fuel, data=SaratogaHouses)

stripchart(price ~ fuel, data=SaratogaHouses)

barchart(price ~ fuel, data=SaratogaHouses)
```

##### 5.4
```{r}
# 5.4

#highest median price for fuel=gas
highest_median_prices <- SaratogaHouses %>% 
  group_by(fuel) %>% 
  summarize(median_price = median(price))

#view
highest_median_prices %>% pander()
```

<!-- END: Q5 -->





<!-- START: Q6 -->
## Q6

##### 6.1
```{r}
# 6.1

#plot
xyplot(len ~ dose, groups=supp, data=ToothGrowth, type=c("p","a"), auto.key=TRUE)
```

##### 6.2
```{r}
# 6.2

```

##### 6.3
```{r}
# 6.3

```

##### 6.4
```{r}
# 6.4

```

##### 6.5
```{r}
# 6.5

```

##### 6.6
```{r}
# 6.6

```

##### 6.7
```{r}
# 6.7

```

##### 6.8
```{r}
# 6.8

```

##### 6.9
```{r}
# 6.9

```

##### 6.10
```{r}
# 6.10

```

<!-- END: Q6 -->





<!-- END: Main Tabset -->

<!-- END: ALL -->