---
title: "t-Tests"
subtitle: "Assessment Quiz"
author: "Andrew Seaman"
date: "2025-Feb-1"
output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load libraries
library(mosaic)
library(tidyverse)
library(DT)
library(pander)

df <- Births78
```

Ensure you have library(mosaic) loaded in RStudio and can View(Births78). This dataset records data for every day of the year. In other words, each row of this dataset repesents a day of the year.

## Q1.
> In a typical year, there are 52 weeks. However, 52 x 7 = 364, and as most of us know, there are 365 days in a year. This means that every year, at least one day gets to happen more than 52 times.
</BR></BR>
> Use appropriate R commands and the Births78 dataset to determine which day of the week in 1978 occurred 53 times.

```{r}
# a.
df %>%
  count(wday) %>% 
  pander()
```

## Q2. 

> Let μ represent the average number of births that a occur on a given day of the week, like Wendesday or Thursday.
</BR></BR>
> Use the Births78 dataset in RStudio to test the following hypotheses:
</BR></BR>
H0:μWednesday−μThursday=0
</BR>
Ha:μWednesday−μThursday≠0

```{r}
# a. List of days to test
days <- c("Wed", "Thu")

# b. Subset the data by day
Wednesday <- df$births[df$wday == days[1]]
Thursday <- df$births[df$wday == days[2]]

# c. Run independent samples t test
output <- t.test(Wednesday, Thursday, mu = 0, alternative = "two.side", conf.level = 0.95)

# d. Clean output of test results
pander(output)
```

## Q3. 

>_What two things are required to compute a p-value?_

A test statistic and a sampling distribution of the test statistic.



<!-- END QUIZ -->