---
title: "t-Tests"
subtitle: "Class Activity"
author: "Andrew Seaman"
date: "2025-May-14"
output: 
  html_document:
    css: ../myWork.css
    theme: journal
    toc: true
    toc_float: true
    code_folding: hide
---

<!-- ------ -->
<!-- SETUP: -->
<!-- ------ -->

<!-- Format HTML Output -->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Libraries -->
```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load
library(mosaic)
library(pander)
library(DT)
library(tidyverse)
```

<!-- Data -->
```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load data
KidsFeet = KidsFeet
```

<!-- Setup End -->










<!-- ---------- -->
<!-- QUESTIONS: -->
<!-- ---------- -->
# Questions:

---

## Q1.
> _Explain to a peer what information you can find under the "Overview", "R Instructions", and "Explanation" tabs for each of the three t-tests on the <a href="../../tTests.html">t-Tests</a> page._

<u>t-Tests:</u>

1. __One Sample t Test__
2. __Paired Samples t Test__
3. __Independent Samples t Test__

<u>Tab content description:</u>

* __Overview:__ Questions, requirements, and hypotheses.
* __R Instructions:__ Test function and parameters.
* __Explanation:__ Summary of how to interpret test results.
    
<!-- Q1. End -->

---

</BR>










## Q2.
> _Use the R Instructions to perform a <u>One Sample t-Test</u> of the following hypotheses using the KidsFeet data set. You should get a test statistic of t = -15.532._

<!-- Hypothesis -->
$$
H_0 : \mu_{\text{length of kids' feet}} = 28 \text{ cm}
$$
<!-- Null Hypothesis -->
$$
H_a : \mu_{\text{length of kids' feet}} \neq 28 cm
$$

Yes, I can. Below are the results of a One Sample t test

<!-- t-test -->
```{r}
# a. One sample t test
pander(t.test(KidsFeet$length, mu = 28, alternative="two.sided", conf.level=.95))
```

<!-- Q2. End -->

---

</BR>










## Q3.
> _Use the R Instructions to perform a <u>Paired-Samples t-Test</u> of the following hypotheses using the KidsFeet data set. You should get a test statistic of t = -11.529._

<!-- Hypothesis -->
$$
H_0: \mu_{\text{length} - 3 \cdot \text{width}} = 0
$$
<!-- Null Hypothesis -->
$$
H_a: \mu_{\text{length} - 3 \cdot \text{width}} \neq 0
$$

Yes, I can. Below are the results of a Paired Samples t Test.
<!-- t-test -->
```{r}
# a. Paired samples t test
pander(t.test(KidsFeet$length - 3 * KidsFeet$width, mu = 0))
```

<!-- Q3. End -->

---

</BR>










## Q4.
> _Use the R Instructions to perform an <u>Independent Samples t-Test</u> of the following hypotheses using the KidsFeet data set. You should get a test statistic of t = 1.9174._

<!-- Hypothesis -->
$$
H_0: \mu_\text{Boys foot length} - \mu_\text{Girls foot length} = 0
$$
<!-- Null Hypothesis -->
$$
H_a: \mu_\text{Boys foot length} - \mu_\text{Girls foot length} \neq 0
$$

Yes, I can. Below are the results of a Independent Samples t-Test.

```{r}
# a. Independent samples t test (assume variance)
pander(t.test(length ~ sex, data = KidsFeet, var.equal = FALSE))
```

<!-- Q4. End -->

---

</BR>










## Q5.
> _Discuss with a peer why we create Q-Q Plots and check sample sizes when performing t Tests._

There are certain requirements that must be met to use a t test.

1. __One Sample t Test__
    * The sample is representative of the population. (Having a simple random sample is the best way to do this.)
    * The sampling distribution of the sample mean 𝑥¯ can be assumed to be normal. This is a safe assumption when either (a) the population data can be assumed to be normally distributed using a Q-Q Plot or (b) the size of the sample (n) that was taken from the population is large (at least n > 30, but “large” really depends on how badly the data is skewed).

2. __Paired Samples t Test__
    * The sample of differences is representative of the population differences.
    * The sampling distribution of the sample mean of the differences 𝑑¯(𝑥¯ of the differences) can be assumed to be normal. (This second requirement can be assumed to be satisfied when (a) the differences themselves can be assumed to be normal from a Q-Q Plot, or (b) when the sample size 𝑛of the differences is large.)

3. __Independent Samples t Test__
    * Both samples are representative of the population. (Simple random samples are the best way to do this.)

    * The sampling distribution of the difference of the sample means (𝑥¯1−𝑥¯2) can be assumed to be normal. (This is a safe assumption when the sample size of each group is 30 or greater or when the population data from each group can be assumed to be normal with a Q-Q Plot.)

<!-- Q5. End -->

---

</BR>










## Q6.
> _Discuss with a peer what an appropriate graphic would be for each of the three t tests you performed above._

### {.tabset .tabset-pills .tabset-fade}

<!-- 1. One Sample t Test -->
#### One Sample t-Test

* *Best graphic:* Histogram or dotplot with reference line (population mean)

<!-- Hypothesis -->
$$
H_0 : \mu_{\text{length of kids' feet}} = 28 \text{ cm}
$$
<!-- Null Hypothesis -->
$$
H_a : \mu_{\text{length of kids' feet}} \neq 28 cm
$$

<!-- Histogram -->
```{r}
(
  ggplot(KidsFeet)
    + geom_histogram(aes(x=length), fill="gray", color="black", binwidth=.6)
    + geom_vline(xintercept=28, color="red",  linetype="dashed", linewidth=2)
    + geom_vline(xintercept=mean(KidsFeet$length), color="forestgreen", linetype="dashed", linewidth=2)
    + labs(
        title="Histogram of foot length",
        subtitle="One Sample t-Test (bin = .6cm)",
        caption="(µ = 28cm)",
        x="Length (cm)",
        y="Count"
    )
    + theme_classic()
    + theme(
        plot.title = element_text(hjust=.5, face="bold"),
        plot.subtitle = element_text(hjust=.5, face="bold", color="darkgray"),
        plot.caption = element_text(color="red", size=12)
    ) 
)
```

<!-- QQ Plot -->
```{r message=FALSE, warning=FALSE}
# a. QQ plot for One Sample t Test

# Compute theoretical quantiles & confidence band manually
qq_data <- KidsFeet %>%
  mutate(sample_quantiles = sort(length),
         theoretical_quantiles = qnorm(ppoints(length), mean(length), sd(length))) %>%
  arrange(theoretical_quantiles)

# Plot QQ plot with shaded confidence band
ggplot(qq_data, aes(x = theoretical_quantiles, y = sample_quantiles)) +
  geom_point() +  # QQ points
  geom_smooth(method = "lm", se = TRUE, color = "black", fill = "forestgreen") +  # Shaded confidence band
  labs(title = "QQ Plot of Foot Lengths in KidsFeet Dataset",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  theme_minimal()
```



<!-- 2. Paired Samples t Test -->
#### Paired Samples t-Test

* *Best graphic:* Dot-and-line plot (paired differences) or histogram of the differences

<!-- Hypothesis -->
$$
H_0: \mu_{\text{length} - 3 \cdot \text{width}} = 0
$$
<!-- Null Hypothesis -->
$$
H_a: \mu_{\text{length} - 3 \cdot \text{width}} \neq 0
$$

<!-- Histogram of the differences -->
```{r}
# Data: add calculated column
kf = KidsFeet %>% 
  mutate(
    "width3" = width * 3,
    "diff" = round(length - width3, 2)
  )

# Plot: Histogram of the differences
(
  ggplot(kf)
    + geom_histogram(aes(x=diff), binwidth=.7, color="black", fill="gray")
    + geom_vline(xintercept=mean(kf$diff), color="forestgreen", linetype="dashed", linewidth=2)
    + geom_vline(xintercept=0, color="red", linetype="dashed", linewidth=2)
    + labs(
        title="Histogram of the Differences ",
        subtitle="Two Sample t-Test (bin = .7cm)",
        caption="(µ = 0cm)",
        x="Difference (cm)\n(Length - (Width * 3))",
        y="Count"
    )
    + theme_classic()
    + theme(
        plot.title = element_text(hjust=.5, face="bold"),
        plot.subtitle = element_text(hjust=.5, face="bold", color="darkgray"),
        plot.caption = element_text(color="red", size=12)
    ) 
)
```

<!--  Dot-and-line plot -->
```{r}
# Convert the kf dataframe to long format
kf_long <- kf %>% 
  mutate( id = row_number()) %>% 
  select(id, length, width3) %>% 
  pivot_longer(cols=c(length, width3), names_to = "condition", values_to = "value")

# Plot: Dot-and-line plot
(
  ggplot(kf_long, aes(x=condition, y=value, group=id, color=condition))
    + geom_point(size=2)
    + geom_line(color="gray")
    
    + labs(
        title = "Dot-and-Line Plot of Paired Differences",
        subtitle = "Paired Sample t-Test",
        caption = "(width3 = width * 3)",
        x = "Condition",
        y = "Measurement (cm)"
      )
    + theme_light()
    + theme(
        plot.title = element_text(hjust=.5, face="bold"),
        plot.subtitle = element_text(hjust=.5, color = "darkgray"),
        plot.caption = element_text(color = "darkgray"),
        legend.position = "none"
      )
)


```



<!-- 3. Independent Samples t Test -->
#### Independent Samples t-Test

* *Best graphic:* Boxplot or side-by-side dotplot / violin plot

<!-- Hypothesis -->
$$
H_0: \mu_\text{Boys foot length} - \mu_\text{Girls foot length} = 0
$$
<!-- Null Hypothesis -->
$$
H_a: \mu_\text{Boys foot length} - \mu_\text{Girls foot length} \neq 0
$$

<!-- Box plot -->
```{r}
# Colors
colors_sex <- c("royalblue", "hotpink")

# Plot: Box plot
(
  ggplot(KidsFeet, aes(x=sex, y=length))
    + geom_boxplot(fill=colors_sex)
    + scale_x_discrete(labels= c("B"="Boy", "G"="Girl"))
    + theme_gray()
    + labs(
        title="Foot Length Comparison by Sex",
        subtitle="Independent Samples t-Test",
        x="Sex",
        y="Length (cm)"
      )
    + theme(
        plot.title = element_text(hjust=.5, face="bold"),
        plot.subtitle = element_text(hjust=.5, color = "darkgray")
      )
)
```

###

<!-- Q6. End -->

---

</BR>










## Q7.
> _Discuss with a peer what insights can be made about each of the three t-tests you performed above. _

Below are the three t-tests and one relevant insight per.

```{r}
insights_df <- data.frame(
  "t-Test" = c("One Sample", "Paired Samples", "Independent Samples"),
  Insight = c(
    "The sample mean was closer to 25cm than 28cm.",
    "The differences on average were just above 2cm off of 0 in the negative direction.",
    "The difference in length by gender is not 0.")
)

pander(insights_df, justify="left", style="grid")
```

<!-- Q7. End -->


<!-- -------- -->
<!-- FILE END -->
<!-- -------- -->
