---
title: "Kruskal-Wallis Test"
subtitle: "Class Activity"
author: "Andrew Seaman"
date: "2025-MM-DD"
output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
    code_folding: hide
---

<!-- ------ -->
<!-- SETUP: -->
<!-- ------ -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load libraries
library(mosaic)
library(DT)
library(pander)
library(tidyverse)
library(car)
```

<!-- Setup End -->










<!-- ---------- -->
<!-- QUESTIONS: -->
<!-- ---------- -->

## Q1.
> _Open your Statistics Notebook to the "Kruskal-Wallis Test" page._

Done.

<!-- Q1. End -->





## Q2.
> _Open the "Explanation" tab from the "Kruskal-Wallis Test" page and discuss with your neighbor any similarities this test has to the Wilcoxon Rank Sum test. Also discuss any differences._

### Similarities
```{r}
similarities <- data.frame(
  Feature = c(
    "Nonparametric",
    "Rank-based",
    "Test for differences between groups",
    "Used when assumptions for ANOVA/t-test are violated"
  ),
  `Wilcoxon Rank Sum & Kruskal-Wallis` = rep("✅", 4)
)

pander(similarities)
```

⸻

### Differences
```{r}
differences <- data.frame(
  Feature = c(
    "# of groups",
    "Also known as",
    "Test statistic",
    "Post-hoc test needed?",
    "Pairing"
  ),
  `Wilcoxon Rank Sum` = c(
    "2 groups only",
    "Mann–Whitney U Test",
    "U or W",
    "No (just 2 groups)",
    "Independent samples"
  ),
  `Kruskal-Wallis` = c(
    "3+ groups",
    "One-way ANOVA on ranks",
    "H",
    "Yes (if H is significant, follow up with pairwise comparisons)",
    "Multiple independent groups"
  )
)

pander(differences)
```

<!-- Q2. End -->










## Q3.
> _What can you conclude when the p-value of the Kruskal-Wallis Test is significant?_

When the p-value of the Kruskal-Wallis Test is significant (typically p < 0.05), you can conclude:

⸻

✅ At least one group differs from the others

There is statistical evidence that at least one group’s distribution (usually median) is significantly different from the others.

⸻

🧠 But! You cannot conclude:

* Which specific group(s) are different
* How many groups are different
* The direction of the difference (e.g. higher or lower)

<!-- Q3. End -->










## Q4.
> _What are the hypotheses of the Kruskal-Wallis Test?_

$$
H_0: \text{All group distributions are equal}
$$

$$
H_A: \text{At least one group distribution differs}
$$
<!-- Q4. End -->










## Q5.
> _Use the Friendly dataset from library(car) in R to perform a Kruskal-Wallis Test._

```{r}
# import data
df <- Friendly
```

```{r}
# Kruskal-Wallis Test
df %>% 
  kruskal.test(correct ~ condition, data = .) %>% 
  pander()
```

<!-- Q5. End -->










## Q6.
> _What advantage does the Kruskal-Wallis Test have over the Wilcoxon Test? What advantage does the Wilcoxon Test have over the Kruskal-Wallis Test?_

### Kruskal-Wallis Test
```{r, echo=FALSE, results='asis'}
kw_adv <- data.frame(
  Advantage = c(
    "Handles 3 or more groups",
    "Good for detecting broader distributional differences"
  ),
  Why_it_matters = c(
    "Wilcoxon is limited to comparing 2 groups only. Kruskal-Wallis generalizes that to multiple independent groups.",
    "Useful when testing whether any group differs, not just one pair."
  )
)

pander(kw_adv)
```

⸻

### Wilcoxon Test
```{r, echo=FALSE, results='asis'}
wilcoxon_adv <- data.frame(
  Advantage = c(
    "More power for 2-group comparisons",
    "Supports post-hoc comparisons"
  ),
  Why_it_matters = c(
    "When you're only testing two groups, Wilcoxon is typically more sensitive and powerful than Kruskal-Wallis.",
    "Often used as the follow-up test after Kruskal-Wallis to detect which groups differ."
  )
)

pander(wilcoxon_adv)
```

<!-- Q6. End -->










## Q7.
> _For your Analysis this week, what three R-codes are needed to complete the analysis?_

__We will need:__

1. `kruskal.test()`
2. `boxplot()` from the ggplot package
3. `favstats()` from library Mosaic

<!-- Q7. End -->





<!-- Questions End -->





<!-- -------- -->
<!-- FILE END -->
<!-- -------- -->