---
title: "Kruskal-Wallis Test"
subtitle: "Skills Quiz"
author: "Andrew Seaman"
date: "2025-Jun-27"
output: 
  html_document:
    theme: journal
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load libraries. Modify list as needed.
library(tidyverse)
library(mosaic)
library(mosaicData)
library(DT)
library(pander)
```

<!-- Data -->
```{r}
# Import data
saratoga <- SaratogaHouses
```

<!-- START: ALL -->
# Questions





<!-- START: Main Tabset -->
# {.tabset .tabset-pills .tabset-fade}





<!-- START: Summary -->
## Summary

----

### Introduction
The Kruskal-Wallis Rank Sum Test "allows for deciding if several samples come from the same population or if at least one sample comes from a different population".

<BR />

### Questions [5]

##### Q1
* Run a Kruskal-Wallis Rank Sum Test (`kruskal.test(x ~ g, data = ...)`).
* Run a hypothesis test.
* Conclude the results with a numerical summary.

##### Q2
* Review example analyses.

##### Q3
* Read the Kruskal-Wallis Test page of the textbook.

##### Q4
* Email Brother Saunders about how I've organized the Table of Contents (`index.rmd`) file to suit my needs in navigating the textbook.

##### Q5
* Check for understanding on the material covered in the __Explanation__ tab on the Kruskal-Wallis Test page of the textbook.

<!-- END: Summary -->





<!-- START: Q1 -->
## Q1

----

##### Introduction
Does the home __resale value increase__ more from upgrading to <u>gas</u> or <u>electric</u> fuel systems?

##### 1.1
```{r}
# Columns in use: 
#   1. price
#   2. fuel
saratoga %>% 
  kruskal.test(price ~ fuel, data = .) %>% 
  pander()
```

##### 1.2 {.tabset .tabset-pills .tabset-fade}

###### Boxplot
__Best__ depicts the results of the Kruskall-Wallis Test.
```{r}
saratoga %>% 
boxplot(price ~ fuel, data = .)
```

###### Stripchart 
Does __not__ best depicts the results of the Kruskall-Wallis Test.
```{r}
saratoga %>% 
stripchart(price ~ fuel, data = .)
```

###### Barchart
Does __not__ best depicts the results of the Kruskall-Wallis Test.
```{r}
saratoga %>% 
barchart(price ~ fuel, data = .)
```

<!-- End tabset of charts -->
#####

##### 1.3
```{r}
# Highest median (price) of the three fuel types? (answer: gas at $206,500)
saratoga %>% 
  favstats(price ~ fuel, data = .) %>% 
  pander()
```

<!-- END: Q1 -->





<!-- START: Q2 -->
## Q2

----

##### 2.1
Review the two example analyses (wages, pigweights) from the instruction page on the Kruskal-Wallis Rank Sum Test and provide the requested test results and hypothesis test results.

* Done

<!-- END: Q2 -->





<!-- START: Q3 -->
## Q3

----

##### 3.1
On the Kruskal-Wallis Test page, review the material for each tab and section.

* Done
<!-- END: Q3 -->





<!-- START: Q4 -->
## Q4

----

##### 4.1

Email your instructor a single screen shot showing the work you have done on your Table of Contents page so far.

* I have emailed my instructor a screen shot of my "Table of Contents" page of my Math 325 Notebook. It shows some of what I have been doing to customize this page.

<!-- END: Q4 -->





<!-- START: Q5 -->
## Q5

These questions test your understanding of the Explanation tab for the Kruskal-Wallis Test page of your Math 325 Notebook.

----

##### 5.1
How many steps are there to calculating a p-value for the Kruskal-Wallis Test?

* 6 steps

<!-- END: Q5 -->






#

<!-- END: Main Tabset -->

<!-- END: ALL -->