---
title: "Wilcoxon Tests"
subtitle: "Assesment Quiz"
author: "Andrew Seaman"
date: "2025-06-04"
output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
    code_folding: hide
---

<!-- ------ -->
<!-- SETUP: -->
<!-- ------ -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load libraries
library(mosaic)
library(DT)
library(pander)
library(tidyverse)
library(car)
```

<!-- Setup End -->










<!-- ---------- -->
<!-- QUESTIONS: -->
<!-- ---------- -->

## Q1.
> _Use the Salaries dataset in R to find the number of male and female assistant professors in the dataset._

```{r}
# a.
Salaries <- Salaries

# b.
Salaries2 <- Salaries %>% 
  filter(rank == "AsstProf")

# c.
count(Salaries2, sex)
```

<!-- Q1. End -->










## Q2.
> _Create an appropriate graphic using the Salaries dataset in R that would allow you to compare the distribution of salaries for faculty in discipline A ("theoretical") and discipline B ("applied") departments._

```{r}
# a.
ggplot(Salaries, aes(x=discipline, y = salary))+
  geom_boxplot() +
  scale_y_continuous(
    breaks = c(100000, 150000, 200000)
  ) +
  labs(
    title = "Faculty From U.S. Colleges\n",
    x = "Discipline",
    y = "Salaries"
  ) +
  theme_bw() +
  theme(
    plot.title = element_text(face = "bold", hjust = .5),
    axis.text.y = element_text(angle = 90),
    grid
  )
```

<!-- Q2. End -->












## Q3.
> _Perform an appropriate Wilcoxon Test of the following hypotheses using the Salaries dataset in R._

$$
H0: {\text{median discipline A salaries - median discipline B salaries}} = 0
\\
Ha: {\text{median discipline A salaries - median discipline B salaries}} \neq 0
$$

Select the response below showing the correct test statistic, p-value, and conclusion of the test based on a significance level of 0.05.

```{r}
# a.
wilcox.test(Salaries$salary[Salaries$discipline == "A"], 
            Salaries$salary[Salaries$discipline == "B"], 
            mu = 0, 
            alternative = "two.sided", 
            conf.level = 0.05)
```
<!-- Q3. End -->





<!-- Questions End -->





<!-- -------- -->
<!-- FILE END -->
<!-- -------- -->