---
title: "Wilcoxon Tests, Boxplots, and P-Values"
subtitle: "Class Activity"
author: "Andrew Seaman"
date: "2025-5-21"
output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
    code_folding: hide
---

<!-- ------ -->
<!-- SETUP: -->
<!-- ------ -->

<!-- Format HTML Output -->
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Libraries -->
```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load libraries
library(mosaic)
library(DT)
library(pander)
library(tidyverse)
```

<!-- Setup End -->




<!-- Tabset Start -->
# {.tabset .tabset-pills .tabset-fade}

<!-- ---------- -->
<!-- QUESTIONS: -->
<!-- ---------- -->

## Questions





### Q1.
> _Let's review some of the things we have learned so far this semester:_

#### Response

1. __Which function would you use to add a new column to your dataset?__
    * We would use `mutate()` or `summarize()`.

```{r}
KidsFeet2 <- KidsFeet %>%
  
  mutate(
    birthseason = case_when(
      birthmonth %in% c(12, 1, 2) ~ "Winter",
      birthmonth %in% c(3,4,5) ~ "Spring",
      birthmonth %in% c(6,7,8) ~ "Summer",
      birthmonth %in% c(9,10,11) ~ "Fall"
    )
  )

  datatable(head(KidsFeet2))
```

</BR>

2. __Which plots work best to display <u>one quantitative variable</u> that is separated into 2 or more groups?__
    * Box plots, violin plots, and dot/strip plots work well. They visualize distribution, central tendency, and spread.
    * From Brother Saunders:
      * If the sample size is fairly large, a boxplot could be used.
      * If the sample size is small, a dot plot is best.
      * If you want to get fancy with medium sample sizes, you could overlay a dotplot on top of a boxplot using either Base R:
      
```{r}
boxplot(length ~ sex, data=KidsFeet)

stripchart(length ~ sex, data=KidsFeet, method="stack", vertical=TRUE, add=TRUE)
```

OR ggplot2:

```{r}
ggplot(KidsFeet, aes(x=sex, y=length)) + 

  geom_boxplot() +
  
  geom_point()
```

</BR>

3. __What two things are needed to calculate a p-value?__
    * A test statistic (e.g., t-score)
    * The sampling distribution under the null hypothesis

</BR>

4. __When would an independent samples t test be inappropriate to use but a Wilcoxon Rank Sum Test be appropriate?__
    * When the data are not normally distributed or when the sample sizes are small and unequal with strong outliers — Wilcoxon doesn’t assume normality.
    * (Brother Saunders): The independent samples t test is appropriate when the sample sizes are large or when the data in each group can be considered to be normal. It is inappropriate to use an independent samples t test when the sample sizes are small AND the data is clearly not normal in either group. This latter situation would still be appropriate for a Wilcoxon Rank Sum Test (as long as there weren't too many "ties" in the data). The Wilcoxon Test can also be used on "ordinal" data, but the t test is not meaningful for such data.

</BR>

5. __What is an appropriate set of hypotheses to use for an independent samples t test? For a Wilcoxon Rank Sum Test?__
    * __t Test__:
      * H₀: μ₁ = μ₂ (the means are equal)
      * H₁: μ₁ ≠ μ₂ (the means are different)

    * __Wilcoxon Rank Sum Test__:
      * H₀: The distributions of the two groups are equal
      * H₁: The distributions are not equal (often interpreted as a shift in median)

---

<!-- Q1. End -->










### Q2.
> _Your instructor will use the following applet for today's class activity. You are welcome to follow along if you wish ([Mann-Whitney / Small Samples](https://shiny.byui.edu/StatClass/))._

#### Response

1. __Explain what it means for the null hypothesis of a Wilcoxon Rank Sum Test to be true.__
    * It means the two groups come from the same distribution — there’s no systematic difference in ranks between them (i.e., no shift in medians).

</BR>

2. __Explain what it means for the alternative hypothesis of a Wilcoxon Rank Sum Test to be true.__
    * It means the distributions differ — one group consistently has higher or lower ranks than the other, suggesting a shift in central tendency.

</BR>

3. __Explain how boxplots for small sample sizes can be misleading.__
    * With small samples, boxplots can exaggerate or hide outliers, skewness, and spread. The quartiles might not represent the underlying distribution well, making patterns appear that aren’t statistically meaningful.

</BR>

4. __Explain how the p-value helps us judge whether or not the null hypothesis or alternative hypothesis is the truth based on the evidence shown in a boxplot.__
    * The p-value quantifies how likely it is to see a difference as extreme as the one in the plot if the null hypothesis were true. A small p-value suggests the difference isn’t just a fluke and supports the alternative hypothesis.

---

<!-- Q2. End -->










### Q3.
> _This week's analysis asks you to examine the results from a memory recall study. It is helpful to the success of this analysis if you experience some of the experiment yourself. To do this, open this [link](https://brainconnection.brainhq.com/brain-teasers/word-list-recall). Then, try the "Word List Recall" game and see how you do._

#### Response

1. __How many words did you correctly recall?__
    * I recalled 4 words.

<!-- Q3. End -->

---

<!-- Questions End -->
 









<!-- ------ -->
<!-- NOTES: -->
<!-- ------ -->

## Notes

### Additional Resources

1. [Wilcoxon Tests](../../WilcoxonTests.html)

---

### `Mean()` vs. `Median()`

The __Wilcoxon tests__ are more usefull than t-Tests when the <span style="background-color:yellow;">data cannot be trusted to be normal</span>, so the `mean()` is prone to following the tail of the data whereas the `median()` is far more stationary. 

By using the `median()` we get a better sense for proportions of real or potential data points which belong in the various quantiles.

---

<!-- Notes End -->

#

<!-- Tabset End -->



<!-- -------- -->
<!-- FILE END -->
<!-- -------- -->