---
title: "Wilcoxon Tests"
subtitle: "Class Activity"
author: "Andrew Seaman"
date: "2025-05-19"
output: 
  html_document:
    theme: journal
    toc: true
    toc_float: true
    code_folding: hide
---

<!-- ------ -->
<!-- SETUP: -->
<!-- ------ -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load libraries
library(mosaic)
library(tidyverse)
library(DT)
library(pander)
```

<!-- Setup End -->










<!-- ---------- -->
<!-- QUESTIONS: -->
<!-- ---------- -->

## Q1.
> _Describe how a Wilcoxon Test computes a p-value (summarize the five steps as listed in the Explanation tabs of the Wilcoxon Test page)._

__Summarization of steps:__

1. Order all the data by magnitude ignoring the sign (+/-). Ensure the group label is kept with the value.
2. Rank all values. 
    * _Singed-Rank_: reassign the sign (+/-) to the rank. 
    * _Rank Sum_: for identical values (i.e. ties) average their ranks and assign the result as the rank for each identical value. Do this for every set of repeated values. 
3. Separate the ranks back into their respective groups.
    * _Signed Rank_: negatives and positives
    * _Rank Sum_: Group 1 and Group 2 (i.e. A and B, etc.).
4. Sum the ranks in a group to find the test statistic.
    * _Signed-Rank_: usually the smaller group (fewer ranks).
    * _Rank Sum_: either group.
5. Compute the p-value. 
    a. Compute the probability of all possible values the test statistic could have (mathematical counting techniques).
    b. Computing the probabilites of all sums creates a distribution of the test statistic. 
    c. Then you can compute the p-value and reject of fail to reject the null hypothesis.

<!-- Q1. End -->










## Q2.
> _Perform the following Wilcoxon Test in R: `wilcox.test(length ~ sex, data=KidsFeet)`._

```{r, warning=FALSE}
# a. Perform the following Wilcoxon test (Mann-Whitney; i.e. equivalent of independent samples t test) in R.
pander(wilcox.test(length ~ sex, data=KidsFeet))

# b. Locate W and the p-value in the output of this test in R. 
W <- 252
p_value <- 0.0836
```
> _Locate W and the p-value in the output of this test in R._

In this Wilcoxon test:

* __W (test statistic)__: `r W`
* __p-value__: `r p_value`

> _Explain what `W` represents and what the p-value represents with regards to the five steps of a Wilcoxon Test._

The `W` represents the test statistic computed through the first four steps of each test.

<!-- Q2. End -->










## Q3.
> _From the output in R, explain what this sentence means: "alternative hypothesis: true location shift is not equal to 0."_

$$
H_a: {\text{true location shift}} \neq 0
$$

The "location shift" refers to the difference in central tendency (typically the difference of medians).

Therefore, the sentence can be rephrased as, "We're testing if the two groups are different (not equal)."

In conclusion, in this case according to the results (p-value = `r p_value`) we actually must _fail_ to reject the null hypothesis that the "location shift" (difference of medians) is equal to 0.

<!-- Q3. End -->










## Q4.
> _Explain the various types of hypotheses that can be used (null and alternative) when performing a Wilcoxon Test. How do these hypotheses differ from those of a t Test?_

Below are the two types of Wilcoxon test and the different sets of null and alternative hypotheses one can test based on the type and orignal of data involved. 

The below sets of hypotheses differ from those used in t tests in at least three ways:

1. The _median_ is used as the measure of central tendency instead of the _mean_.
2. Wilcoxon tests and hypotheses are used to evaluate nonparametric distributions.
3. The results of Wilcoxon test and nonparametric tests at large are less powerful, or less convincing, than t tests in terms of the results and conclusions.

### Wilcoxon Signed-Rank Test

__One Sample of Data__
$$
  H_0: \text{Median} = \text{(Some Number)}
    \\
  H_a: \text{Median} \neq \text{(Same Number)}
$$

__Paired Samples of Data__
$$
  H_0: \text{median of differences} = 0
    \\
  H_a: \text{median of differences} \neq 0
$$


### WIlcoxon Rank Sum Test

__Independent Samples of Data__
$$
  H_0: \text{difference in medians} = 0
    \\
  H_a: \text{difference in medians} \neq 0
$$

Or more generally,"that one distrubution is <u>_stochastically greater_</u> than the other."
$$
  H_0: {\text{the distributions are stochastically equal}}
    \\
  H_a: {\text{one distribution is stochastically greater than the other}}
$$

<!-- Q4. End -->










## Q5.
> _Explain what this error message means: "cannot compute exact p-value with ties" and what the implications are when this message appears while performing a Wilcoxon Test._

__Error message__

* The Wilcoxon Rank Sum Test (and Signed Rank Test) is based on ranks.
* The test assumes that all values can be uniquely ranked — meaning no ties (no identical values).
* If there are ties (duplicate values), the algorithm can’t compute an exact p-value (which is based on all possible permutations of the ranks).
  * _Why that blocks “exact p-value”_:
    * The exact distribution of the test statistic relies on counting all possible distinct arrangements of ranks.
    * Ties create __ambiguity__ → the exact enumeration formula doesn’t apply anymore → R falls back to an approximation.

__It's implications__

* R switches to an __approximate__ p-value: usually a __normal approximation__ (with or without continuity correction).
* The result is still __valid__: but less precise than an "exact" sample result.
* Common when data has __rounding__ or __discrete values__: ex. Likert scales, counts, rounded measures.

<!-- Q5. End -->



<!-- Questions End -->



<!-- -------- -->
<!-- FILE END -->
<!-- -------- -->