---
title: "Chi Squared Test"
subtitle: "Skills Quiz"
author: "Andrew Seaman"
date: "2026-04-09"
output: 
  html_document:
    theme: simplex
    code_folding: hide
---

<!-- IMPORTS -->
```{r, message=FALSE, warning=FALSE, include=FALSE}
# Load libraries
pacman::p_load(mosaic, DT, pander, tidyverse)
```

<!-- START: Main Tabset -->

# Pages 

# {.tabset .tabset-pills .tabset-fade}





<!-- START: Summary -->

## Home


### Introduction
The focus of this Skills Quiz is to learn how to setup, perform, and interpret a Chi-Squared based analysis.


### Questions 

##### Q1
Introduction and textbook study.

##### Q2
Fill in the blank.

##### Q3
\[summary]

##### Q4
\[summary]

##### Q5
\[summary]

<!-- END: Summary -->





<!-- START: Q1 -->
## Q1

##### 1.1
Mathematical model for a basic Chi-Squared test:
$$
\chi^2 = \sum_{i=1}^m \frac{(O_i - E_i)^2}{E_i}
$$

<!-- END: Q1 -->





<!-- START: Q2 -->
## Q2

Fill in the blank.

<!-- END: Q2 -->





<!-- START: Q3 -->
## Q3

##### 3.1

Run the following code in R. (This data is fabricated to mimic reality, but is not actual data.)
 
```{r}
# 3.1 

# Table:

#create
glasses <- cbind( Males = c(Glasses = 5, Contacts = 12, None = 18), Females = c(Glasses = 4, Contacts = 14, None = 22))

#view
glasses %>% 

#plot
barplot(glasses, beside=TRUE, legend.text=TRUE, args.legend=list(x = "topleft", bty="n"))
```

##### 3.2
Run the following code in R to determine if a chi-squared test of independence is appropriate for the glasses data.
```{r, warning=FALSE}
# 3.2

#chi-squared test
chis.glasses <- chisq.test(glasses)

#expected values
chis.glasses$expected %>% pander()
```

##### 3.3
Which of the following statements are correct interpretations of the bar plot that the above code generated in R?
```{r}
# 3.3

mean(4.20000, 4.80000, 12.13333, 13.86667, 18.66667, 21.33333)
```

##### 3.4
Run the following code in R to view the results of your test.
```{r}
# 3.4

chis.glasses %>% pander()
```

##### 3.5
Test residuals:
```{r}
# 3.5

chis.glasses$residuals  %>% pander()
```

<!-- END: Q3 -->





<!-- START: Q4 -->
## Q4

##### 4.1

The following code gets this data into R.

```{r}
# 4.1

#build
education <- cbind( `United States` = c(Engineering = 61941, `Natural Science` = 111158, `Social Science` = 182166), `Western Europe` = c(Engineering = 158931, `Natural Science` = 140126, `Social Science` = 116353), Asia = c(280772, 242879, 236018))

#view
education %>% pander()
```

##### 4.2
```{r}
# 4.2

#chi-squared test
chis.education <- chisq.test(education)

#view
chis.education %>% pander()

#plot
barplot(education, beside=TRUE, legend.text=TRUE, args.legend=list(x = "topleft", bty="n"))
```

##### 4.3
Pearson test residuals
```{r}
# 4.3

#residuals
chis.education$residuals  %>% pander()
```

<!-- END: Q4 -->





<!-- START: Q5 -->
## Q5

##### 5.1
Consider the InsectSprays dataset in R. 

\> ?InsectSprays

Test the hypothesis that the mean number of bugs killed by insectisides A, B, C, D, E, and F are all the same against the alternative that at least one mean is different.
```{r}
# 5.1

#contingency table
sprays <- table(InsectSprays$spray)
sprays %>% pander()

#chi-squared test
chis.sprays <- chisq.test(sprays)
chis.sprays %>% pander()

#expected values
chis.sprays$expected %>% pander()
```

##### 5.2
```{r}
# 5.2

#plot
barplot(sprays, beside=TRUE, legend.text=TRUE, args.legend=list(x = "topleft", bty="n"))
```

##### 5.3
Apparently I'm supposed to find the f-statistic for this test because ANOVA is a better fit for the data than chi-squared.
```{r}
# 5.3

#anova
anova_model <- aov(count ~ spray, data = InsectSprays)
summary(anova_model)
```

<!-- END: Q5 -->





<!-- END: Main Tabset -->





<!-- END: ALL -->