---
title: "Skills-Quiz_Intro-to-R"
author: "Andrew Seaman"
date: "2025-01-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
"Note: used in number 3."
library(ggplot2)
library(plotly)
```

# Mobius - Skills Quiz - Introduction to R

## R help -> R Commands 

### $ The Selection Operator
```{r}
#1

"a"
airquality$Wind

"b"
m <- mean(airquality$Wind)
round(m, digits = 3)

"c"
s <- sd(airquality$Wind)
round(s, digits = 3)

"d"
m <- mean(airquality$Temp)
round(m, digits = 1)
"(output: 77.9)"
```

### <- The Assignment Operator
```{r}
#2

"a"
cars2 <- cars

"b"
cars2$ftpersec <- cars2$speed * 5280 / 3600

"c"
View(cars2)

"d"
airquality2 <- airquality
airquality2$Celcius <- (airquality$Temp - 32) * (5/9)
View(airquality2)
```

### Scatterplots
```{r}
#3.1 (Base R)

"a"
plot(airquality$Temp, type="l")

"b"
plot(Temp ~ Wind, data=airquality, pch=21)

"Note: you can insert an ifelse statement to conditionalize the 'pch' here."
plot(Temp ~ Wind, data=airquality, pch=ifelse(airquality$Temp >= 75, 24, 25))

"c"
plot(Temp ~ Wind, data=airquality, 
     xlab="Daily Wind Speed (mph)",
     ylab="Temperature",
     main="La guardia Airport (May - Sep)",
     col="ivory3", pch=18
     )
```

```{r}
#3.2 (ggplot2)

"a"
ggplot(airquality, aes(x=Wind, y=Temp)) +
  geom_point()

"b"
ggplot(airquality, aes(x=Wind, y=Temp)) +
  geom_point(color = "ivory3", pch = 18) +
  labs(
    title="La Guardia Airport (May - Sep)",
    x="Daily Average Wind Speed (mph)",
    y="Daily Mean Temperature") +
  theme_bw()
```

```{r}
#3.3 (plotly)

"a"
plot_ly(
  data = airquality,
  x = ~Wind,
  y = ~Temp,
  type = 'scatter',
  mode = 'markers'
)

"b"
"Note: 'KidsFeet' is not a dataset found in Base R. Unclear where it comes from."
plot_ly(KidsFeet,
        x= ~length,
        y= ~width,
        color= ~sex,
        size= ~ birthmother,
        text= ~paste("Name:", name, "\n", "Birth-Month:", birthmonth),
        colors=c("skyblue","hotpink")) %>%
  layout(title="KidsFeet dataset",
         xaxis=list(title="Length of the longer foot in cm"),
         yaxis=list(title="Width of the longer foot in cm"))
```

```{r}
#3.4 (Prove you understanding)

"a."
plot1 <- ggplot(airquality, aes(x=Temp, y=Ozone)) +
  geom_point(color = "firebrick", pch=16) +
  labs(
    title="Exponential Growth in Ozone with Increasing Temp",
    x="Daily Maximum Temperature",
    y="Mean Ozone in Parts per Billion (from 1300 to 1500 hours)"
  ) +
  # Note: The following was developed with the help of Chat GPT.
  theme_minimal(base_size = 15) +
  theme(
    panel.background = element_rect(fill = "white", color = "black"),
    panel.grid = element_blank(),
    axis.ticks.length = unit(0.3, "cm"),
    axis.ticks = element_line(color = "black", linewidth = 0.5),
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 13, face = "bold", hjust = 0.5),
    axis.title.x = element_text(size = 9),
    axis.title.y = element_text(size = 7),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 10, 10,)
  )

plot1

```

### Dot Plots
```{r}
#1

"a"
ggplot(airquality, aes(x=Temp)) +
  geom_dotplot()

"b"
ggplot(airquality, aes(x=factor(Month), y=Temp)) +
  geom_dotplot(binaxis = "y", stackdir = "up", position = "dodge", dotsize = 0.75, binwidth = 0.5)

"c"
ggplot(airquality, aes(x=factor(Month), y=Temp)) +
  coord_flip() +
  geom_dotplot(binaxis = "y", stackdir = "up", position = "dodge", dotsize = 0.25, bindwidth = 0.5) +
  labs(title="La Guardia Airport Daily Mean Temperature", x="Month of the Year", y="Daily Mean Temperature")

"d" 
ggplot(airquality, aes(x=factor(Month), y=Ozone)) +
  coord_flip() +
  geom_dotplot(binaxis = "y", stackdir = "up", position = "dodge", dotsize = 0.25, bindwidth = 0.5) +
  labs(title="La Guardia Airport Daily Mean Ozone", x="Month of the Year", y="Mean Ozone in Parts per Billion (from 1300 to 1500 hours)")
```
